### 2012-02-24 papers/eccv2012/main.tex
sections 2 and 3 written

\begin{abstract}
  We consdier the problem of learning to reconstruct polygonal models
  from single and multiple views.

  In this difficult domain we start small and focus on indoor
  Manhattan models.

  We learn lots about practical structured prediction in this unique
  domain

  We kick the state-of-the-art's butt on everything :)
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Reconstruction long considered a central problem in computer vision

Multiple view reconstruction long history, mostly uses probabilistic
models with few free parameters

Single view reconstruction a recent but now well-established
trend. Learning obviously a core part of this fundamentally
under-determined problem. But not much care on loss functions,
optimality guarantees, or statistical consistency.
-> Exceptions: Hedau et al (but very simple models)

Structured prediction used for increasingly complex models. Its use in
computer vision part of a long trend towards statistically rigorous
and well-understood convex optimization techniques for solving
empirical risk minimization problems in computer vision.

Has been used much before for reconstruction. Stereo for example, and
Hedau. Here we go all the way to an expressive high-level geometric
model class.

Attraction of doing this is many fold:
Can minimize model-level loss rather than pixel-level loss
-> give a nice demonstration of this
Can minimize realistic loss functions in a rigorous way
Can learn in both single and multiple view setting using the *same*
model
Bayesian inference algorithm

We learn much about the practice of structured prediction of value to
the community and focus an entire section of conveying these lessons.

Out approach: describe the hypothesis class, describe the model

The remainder of this paper is...

\section{Background}
\label{sec:background}

We're drawing together multiple view reconstruction, single view
reconstruction, and structured prediction -> cannot give a reasonable
survey, sorry!!

Reconstruction papers

Multi view reconstruction, stereo -> not much learning

Single view -> learning but mostly at pixel level. Some exceptions

Learning papers

Structured prediction, reconstruction applications

\section{Model}
\label{sec:model}

We now describe the indoor Manhattan model that we are attempting to
learn. As with any learning problem, there are three ingredients that
we need to specify: a hypothesis class, an observation model for
hypotheses, and a loss function. In our setup these are, respectively,
the class of indoor Manhattan models, a log--linear Bayesian
likelihood, and either the relative depth error or a labelling error
(we describe both).

\subsection{Hypothesis Class}

This paper is concerned with the hypothesis class consisting of indoor
Manhattan reconstructions, which are 3D polygonal models characterized
by infinite floor and ceiling planes with vertical walls extending
between them \cite{Lee09,Flint10cvpr}. Indoor Manhattan environments
are a sub--class of general Manhattan environments. Examples are shown
in figure XXX.

This is an attractive hypothesis class because
\begin{enumerate}
\item{it captures many \textit{regularities} within man--made
  environments;}
\item{the geometric primitives (floor/wall/ceiling) are immediately
  useful for semantic--level scene understanding;}
\item{it is expressive enough to represent approximately or exactly
  a surprisingly wide variety of environments;}
\item{there is a simple and convenient parametrisation;}
\item{an efficient inference algorithm exists \cite{Flint10eccv}.}
\end{enumerate}

An indoor Manhattan model is much like an archicetural floor--plan,
and can be specificied as a set of 2D line segments representing walls
together with the position of the floor and ceiling plane (figure
XXX). In this paper we adopt an image--domain parametrisation due to
its convenience for inference and learning. Following Flint \etal
\cite{Flint11} we represent hypotheses as a path running from the left
edge to the right edge of the image, which we will refer to as a
\textit{seam}. A seam $\Seam$ consists of a sequence of pairs of
scalars,
\begin{equation}
  \Seam = \{\seam_i,\Label_i\}_{i=1}^{\Width}
\end{equation}
where $\seam_i$ is the $y$--cordinate at which the path intersects
image column $i$, $\Orient_i$ is the orientation of the wall in that
column, and $\Width$ is the image width. Remarkably, this simple
parametrisation specifies a unique metric 3D model up to scale
\cite{Flint11}.

While we do not here have space for a full discussion of the geometry
of indoor Manhattan models, there are two properties of the seam
representation that will be of relevance to the following
sections. Firstly, images of indoor Manhattan environments can be
rectified so that vertical lines in the world project to vertical
lines in the image \cite{Flint10eccv}. Secondly, indoor Manhattan
models decompose in the following manner. We said earlier that a seam
$\Seam$ specifies a unique 3D reconstruction. Let
$\Orient(x,y;\Seam)\in\TheOrientations$ be the orientation of the 3D
surface projecting to pixel $(x,y)$. Then Flint \etal
\cite{Flint10eccv} showed that $\Orient$ is functionally dependant
\textit{only} on the labels for column $x$. That is, we may write
\begin{equation}
  \label{eq:orient-indep}
  \Orient(x,y;\Seam) = \tilde{\Orient}(x,y;\seam_x,\Label_x) ~.
\end{equation}
Similarly, letting $\Depth$ be the depth of the reconstruction at
$(x,y)$ we may write $\Depth$ in terms of a single label pair,
\cite{Flint10iccv}
\begin{equation}
  \label{eq:depth-indep}
  \Depth(x,y;\Seam) = \tilde{\Depth}(x,y;\seam_x,\Label_x) ~.
\end{equation}

\subsection{Observation Model}

We adopt the probabilistic model formulated by Flint \etal
\cite{Flint11}, which relates indoor Manhattan hypotheses to
single--view image features, multiple--view photoconsistency terms,
and a reconstructed point cloud. From our perspective these are simply
features and any subset may be omitted if in appropriate for a given
application.

TODO: walk through the graphical model?

Under this model the prior on environments is a geometric distribution
in the number of distinct wall segments, which means, roughly, that we
consider reconstructions with more walls apriori less likely. The
benefit of this prior is that its logarithm is linear in the number of
wall segments,
\begin{equation}
  \log P(\Seam) = - \vect{n} \cdot \Penalties
\end{equation}
where $\vect{n}$ is a vector containing the number of walls in $\Seam$
of various categories (concave, convex, occluding) and $\Penalties$ is
a vector of hyper--parameters.

Flint \etal showed that for a variety of single--view and
multiple--view features $\Features$ there are reasonable choices of
likelihood that can be written in the form
\begin{equation}
  \log P(\Features ~|~ \Seam) = \sum_{x=1}^\Width 
  \PixelPayoff_{\Features}(x,\seam_x)
\end{equation}
where $\PixelPayoff_{\Features}$ is a 2--dimensional matrix computed
deterministically from $\Features$ \cite{Flint11}. It is easy to show
that for each likelihood in \cite{Flint11}, the log--likelihoods
is linear in the hyper--parameters $\LikelihoodHypers$,
\begin{equation}
  \log P(\Features ~|~ \Seam) = 
  \sum_{x=1}^\Width
    \LikelihoodHypers_{\Features} 
    \cdot
    \PixelFtr_{\Features}(x,\seam_x)
\end{equation}

Assuming conditional independence between the feature types, we have
for the posterior,
\begin{eqnarray}
  P(\Seam ~|~ \Features_1 \ldots \Features_n)
  &\propto&
  P(\Seam) P(\Feature_1 ~|~ \Seam) \ldots P(\Feature_n ~|~ \Seam)\\
  \log P(\Seam ~|~ \Features_1 \ldots \Features_n)
  &=&
  - \vect{n} \cdot \Penalties
  + \sum_{i=1}^n \sum_{x=1}^\Width 
      \LikelihoodHypers_i
      \cdot
      \PixelFtr_i(x,\seam_x)
  + O(1) ~.
  \label{eq:log-posterior}
\end{eqnarray}

In order to place our model within a structured prediction framework
we need to re--write the above in terms of a joint feature space
$\JointFtr(\Seam,\Features)$. The decomposability of indoor Manhattan
mdoels into payoff matrices permits precisely such a
formulation. Defining
\begin{align}
  \JointFtr(\Seam,\Features) =
  \begin{bmatrix}
    -\vect{n} \\
    \sum_{x=1}^\Width \PixelFtr_1(x,\seam_x) \\
    \vdots \\
    \sum_{x=1}^\Width \PixelFtr_n(x,\seam_x)
  \end{bmatrix}
  & \quad &
  \Model =
  \begin{bmatrix}
    \Penalties \\
    \LikelihoodHypers_1 \\
    \vdots \\
    \LikelihoodHypers_n
  \end{bmatrix}
  \label{eq:joint-ftr-def}
\end{align}
we see that \eqnref{log-posterior} can be written
\begin{equation}
  \log P(\Seam ~|~ \Features_1 \ldots \Features_n)
  =
  \bigl\langle \Model ~,~ \JointFtr(\Seam,\Features) \bigr\rangle + O(1)
\end{equation}
where we have adopted Hilbert space notation with
$\langle\cdot,\cdot\rangle$ denoting an inner product. Since $\Model$
contains all the free parameters in the model, the goal learning will
be to optimise $\Model$ with respect to a set of training features
$\{\Features_i\}$ together with ground truth labellings $\{\Seam_i\}$.

\subsubsection{Features}

The precise make--up of the feature space depends on the particular
forms for the likelihoods and which sensor modalities are
available. We define separate feature spaces for the single-- and
multiple--view contexts; both are summarized in \figref{featurespace}.

\begin{figure}[tb]
  \centering
  \begin{tabular}{@{}llll@{}}
    \toprule
    Feature & Dimensionality & Single--view? & Reference \\
    \midrule
    Line sweeps & 1 & yes & Lee \etal \cite{Lee09}\\
    Stereo photo--consistency & 4 & no & Flint \etal \cite{Flint11}\\
    Point cloud & 2 & no & Flint \etal \cite{Flint11}\\
    \bottomrule
  \end{tabular}
  \caption{The composition of our feature space. Each row corresponds
    to one or more of the $\PixelFtr_i$ in \eqnref{joint-ftr-def}.}
  \label{fig:featurespace}
\end{figure}

\subsection{Loss Functions}

A loss function $\Delta(\Seam,\EstSeam)$ is a measure of the
deviation of a predicted label $\EstSeam$ from the ground truth
$\Seam$. In the context of learning one often faces a trade--off
between choosing a loss that leads to tractible optimization, and
choosing a loss that measures what one ``really'' cares about. For
example, Hoiem \etal \cite{Hoiem05} implicitly use a per--pixel loss,
since this decomposes optimization pixel--wise. However, it would be
hard to argue that pixel--wise accuracy is an honest measure of what
the authors care about: some pixel--level errors are insignificant to
the overall quality of a reconstruction, while others are
catastrophic. Whether the authors care about the accuracy of their
reconstruction in 3D, its utility for contextual object detection, or
even the aesthetic appeal of reconstructions, a per--pixel loss is
almost certainly a weak measure of success.

A more reasonable choice would be the relative depth error, which has
been the gold standard within the reconstruction community for decades
\cite{Hartley04}, and measures the average deviation between
reconstructed and ground truth depths:
\begin{equation}
  \DepthLoss(\Seam,\EstSeam)
  =
  \frac{1}{N}
  \sum_{\Pixel}
    \frac
        {\Depth(\Pixel;\Seam) - \Depth(\Pixel;\EstSeam)}
        {\Depth(\Pixel;\Seam)} ~,
\end{equation}
where $N$ is the number of pixels. Another reasonable choice is the
labelling error, used widely within the semantic segmentation
literature,
\begin{equation}
  \LblLoss(\Seam,\EstSeam)
  =
  \frac{1}{N}
  \sum_{\Pixel}
    \bigl[~\Orient(\Pixel;\Seam) \neq \Orient(\Pixel;\EstSeam)~\bigr] ~,
\end{equation}
where $[p]$ is 1 if $p$ is true and 0 otherwise. An attractive
characteristic of the indoor Manhattan class is that both of these
loss functions can be optimized exactly. The algorithmic details for
this are left to \sectref{learning}; the key result that we need to
attend to here is that $\DepthLoss$ and $\LblLoss$ can be written in a
form similar to the payoff formulation for the feature
likelihoods. Recall the functional independence established in
\eqnref{func-indep}.

In the
case of reconstruction, the gold standard is the relative depth error
\cite{Hartley04}

Should we show how to write these as payoff matrices here, or leave
this to the learning section?

L1, L2 error: easy representation, not meaningful, give examples

Labelling error
-> showing these will require some of the ICCV framework to have been presented
-> give nice intuition, show scatter plots, examples

Relative depth error
-> showing these will require some of the ICCV framework to have been presented
-> give nice intuition, show scatter plots, examples

None of these are pixel classifiers!

Writing loss functions in the seam representation -> maybe just
mention that this will be useful later

\section{Inference}
\label{sec:inference}

Describe the inference algorithm in terms of payoff/penalty
optimization, cite ECCV -> this section must be short!

\section{Learning}
\label{sec:learning}

Risk minimization formulation
%  -> min_f E[ Loss(f(x),y*) ]
Empirical risk problem
%  -> min_f sum_{xi} Loss(f(x),y*)
Write f as a linear argmax
%  -> f(x) = argmax_y <w,psi(x,y)>
Reduced to quadratic optimization (cite SVMstruct papers)
%  -> min_f ...
-> subject to ....
Minimize using algorithm in ...
This requires a separation oracle, which is the topic of the
following subsection

Loss-augmented inference
-> the problem
-> writing loss the functions as payoff matrices
-> writing a payoff matrix for loss-augmented inference
-> write a proof that ``the optimal model payoff matrix
B=... solves the loss-augmented inference problem <w,psi> + L

\section{Practical Considerations}
\label{sec:practical-considerations}

- negative slacks are more harmful than one would expect

- inconsistent inference is very harmful

- inference when the ground truth not in the hypothesis class

- conditioning in a joint feature space

\section{Multiple View Results}
\label{sec:mv-results}

\section{Single View Results}
\label{sec:sv-results}

\section{Conclusion}
\label{sec:conclusion}

Wow, we really did rock didn't we?

Hypothesis class, why it's interesting

Statistical rigour is guuuud

These fields are ready for this kind of unification

Rigorous joint learning for contextual object detection: a challenge

\bibliographystyle{splncs}
\bibliography{AVLstrings,VisionRefs}

### 2012-03-05 papers/eccv2012/main.tex
ECCV review submission

@@
 \begin{abstract}
-  We consdier the problem of learning to reconstruct polygonal models
-  from single and multiple views.
-
-  In this difficult domain we start small and focus on indoor
-  Manhattan models.
-
-  We learn lots about practical structured prediction in this unique
-  domain
-
-  We kick the state-of-the-art's butt on everything :)
+  We develop a structured prediction approach to reconstructing 3D
+  polygonal models from single and multiple images of a
+  scene. Building on recent advances in single view reconstruction we
+  adopt the indoor Manhattan hypothesis class --- one of the most
+  complicated output spaces (in terms of internal constraints) yet
+  considered within a structured prediction framework. Our approach
+  can learn in both the single-- and multiple--view contexts. We
+  show that the chosen hypothesis class permits optimizing a variety
+  of high--level loss functions, such as the relative depth error. Our
+  results out--perform the state--of--the--art, including an
+  improvement of more than $50\%$ on one metric.
 \end{abstract}
@@
 
-Reconstruction long considered a central problem in computer vision
+Reconstruction 3D models from images is a central problem in computer
+vision. There is a deep literature concerning reconstruction from
+multiple views of a scene, with particular focus on the geometry of
+multiple cameras. More recently the recovery of 3D information from a
+single image has received attention \cite{Hoiem05,Saxena09}. In this
+context the geometric constraints leveraged by multiple--view
+reconstructors are unavailable, so inference must rely on photometric
+cues. To capture the complex, high--dimensional patterns involved in
+this challenging inference task, practitioners have leveraged a range
+of machine learning techniques.
 
-Multiple view reconstruction long history, mostly uses probabilistic
-models with few free parameters
+Within single--view reconstruction, few authors have cast learning as
+a single optimisation with respect to a clearly defined loss function
+\cite{Hoiem05,Saxena09,Lee09,Flint11}, while most approaches to
+multiple--view reconstruction do not consider learning from training
+data at all. In contrast, this paper casts reconstruction
+fundamentally as a learning problem, with the goal being to learn a
+prediction function $\Predictor$ mapping observed features to 3D
+reconstructions.
 
-Single view reconstruction a recent but now well-established
-trend. Learning obviously a core part of this fundamentally
-under-determined problem. But not much care on loss functions,
-optimality guarantees, or statistical consistency.
--> Exceptions: Hedau et al (but very simple models)
+Building on recent advances in single--view reconstruction we adopt as
+our hypothesis class the set of indoor Manhattan models
+\cite{Lee09,Flint10eccv,Flint11}, under which scenes are approximated
+by a floor and ceiling plane together with a sequence of vertical
+walls. This representation brings a variety of attractive features
+such as a simple parametrisation, efficient and exact inference,
+decomposability of loss functions, and a balance between
+expressiveness and robustness.
 
-Structured prediction used for increasingly complex models. Its use in
-computer vision part of a long trend towards statistically rigorous
-and well-understood convex optimization techniques for solving
-empirical risk minimization problems in computer vision.
+For learning we employ the tools of structured prediction, in
+particular the structural SVM \cite{Tsochantaridis04}. The use of
+these tools is part part of a long trend towards statistically
+rigorous, well--understood convex optimization techniques in computer
+vision. Recent successfully applications include detection
+\cite{blaschko2008learning}, segmentation \cite{taskar2004max}, and
+scene classification \cite{Yang2010}. In the domain of reconstruction,
+structured prediction ideas have been applied to several simple models
+classes such as stereo disparities \cite{li2008learning} and cuboids
+\cite{Hedau09}.
 
-Has been used much before for reconstruction. Stereo for example, and
-Hedau. Here we go all the way to an expressive high-level geometric
-model class.
+The application of structured prediction to the indoor Manhattan class
+of models constitutes one of the most complex output spaces yet
+considered within this framework. The indoor Manhattan model enforces
+hard geometric constraints that lack simple expressions in terms of
+image coordinates. These constraints are context--dependent, being
+tied to quantities such as camera rotation and the location of
+vanishing points. We have learnt several valuable lessons of general
+relevance from this complex prediction task, to which we dedicate the
+final section of this paper.
 
-Attraction of doing this is many fold:
-Can minimize model-level loss rather than pixel-level loss
--> give a nice demonstration of this
-Can minimize realistic loss functions in a rigorous way
-Can learn in both single and multiple view setting using the *same*
-model
-Bayesian inference algorithm
+The contributions of this paper are thus (i) a unified learning
+framework for single-- and multiple--view reconstruction, utilising
+the indoor Manhattan model and the tools of structured prediction;
+(ii) the reduction of two image--level loss functions to a form
+amenable to efficient optimization; (iii) an efficient separation
+procedure for identify the ``most--violated constraint'' during
+learning, (iv) an empirical demonstration of structured prediction in
+perhaps the most complex output space yet considered within this
+framework; and (v) a series of practical observations concerning the
+application of structured prediction techniques.
 
-We learn much about the practice of structured prediction of value to
-the community and focus an entire section of conveying these lessons.
-
-Out approach: describe the hypothesis class, describe the model
-
-The remainder of this paper is...
+In the remainder of this paper we present background material
+(\secref{background}), followed by the indoor Manhattan model itself
+(\secref{model}), and our learning framework (\secref{learning}). We
+then present results for multiple--view reconstruction
+(\secref{mv-results}) followed by single--view reconstruction
+(\secref{sv-results}), then we close with practical lessons learnt
+(\secref{discussion}) and concluding remarks (\secref{conclusion}).
 
@@
 
-We're drawing together multiple view reconstruction, single view
-reconstruction, and structured prediction -> cannot give a reasonable
-survey, sorry!!
+This paper touches on several major research areas (multiple--view
+reconstruction, single--view reconstruction, and structured
+prediction), so here we present only key contributions and those
+results of particular relevance to our own work.
 
-Reconstruction papers
+Multiple view reconstruction has a long history in the literature,
+beginning at least as early as the seminal work of Marr \etal
+\cite{Marr1976}. Low--level approaches estimate the depth of each
+pixel by solving for stereo disparity (for a survey see
+\cite{Scharstein2002}), while high--level approaches attempt to
+recover polyhedral models of various kinds
+\cite{Baillard1999,Liebowitz1999}. The present work follows the spirit
+of the latter approach.
 
-Multi view reconstruction, stereo -> not much learning
+Coughlan and Yuille \cite{Coughlan99} first introduced the Manhattan
+world assumption, in which reconstructed surfaces are restricted to three
+mutually--orthogonal orientations. Furukawa \etal \cite{Furukawa09}
+applied this idea within a multiple view stereo context. Lee \etal
+\cite{Lee09} proposed the indoor Manhattan assumption, which places
+further restrictions on reconstructed models.
 
-Single view -> learning but mostly at pixel level. Some exceptions
+Single--view reconstruction from line drawings also has a long history
+in the literature (\cite{Huffman71}, for example). Reconstructing
+single real--world images was re--introduced to the community by Hoiem
+\etal \cite{Hoiem05}, who employed decision trees to classify image
+segments into orientation classes in a multiple segmentation
+approach. Saxena \etal \cite{Saxena09} employed energy minimisation to
+recover pixel--wise depths. Barinova \etal \cite{Barinova08}
+reconstructed piece--wise planar outdoor scenes using an EM
+algorithm. Hedau \etal \cite{Hedau09} modelled indoor scenes as the
+interior of a cuboid, and cast learning within a structured prediction
+framework. We also use structured prediction, though our hypothesis
+class is far more expressive.
 
-Learning papers
+Lee \etal \cite{Lee09} were the first to reconstruct indoor Manhattan
+scenes. They devised a combinatorial search over line segments using a
+branch--and--bound algorithm. Flint \etal
+\cite{Flint10eccv} refined this to an exact dynamic
+programming solution, and in later work \cite{Flint11} extended the
+indoor Manhattan assumption to the multiple view domain. The authors
+of \cite{Flint11} did describe a learning algorithm based on
+bootstrapping, though its statistical consistency was given little
+attention. The present work relies extensively on this thread of
+research, though in contrast to past work our focus is entirely on
+learning in a statistically rigorous framework. We provide comparisons
+with these approaches.
 
-Structured prediction, reconstruction applications
+Extending binary and multi--label classification to general output
+spaces is a major research programme within the machine learning
+community; an excellent introduction is given by Bakir \etal
+\cite{Bakir07}. Our approach employs the structured support vector
+machine, first proposed by Tsochantaridis \etal
+\cite{Tsochantaridis04} and improved upon by Joachims \etal
+\cite{Joachims2009}.
 
@@
 
-We now describe the indoor Manhattan model that we are attempting to
-learn. As with any learning problem, there are three ingredients that
-we need to specify: a hypothesis class, an observation model for
-hypotheses, and a loss function. In our setup these are, respectively,
-the class of indoor Manhattan models, a log--linear Bayesian
-likelihood, and either the relative depth error or a labelling error
-(we describe both).
+In this section we describe three components of the model that we are
+trying to learn (and, at test time, infer): a hypothesis class, a
+feature space, and a loss function. In our setup these are,
+respectively, the class of indoor Manhattan models, a log--linear
+Bayesian likelihood, and either the relative depth error or a
+labelling error (we describe both).
 
@@
 by infinite floor and ceiling planes with vertical walls extending
-between them \cite{Lee09,Flint10cvpr}. Indoor Manhattan environments
-are a sub--class of general Manhattan environments. Examples are shown
-in figure XXX.
-
-This is an attractive hypothesis class because
+between them (as originally proposed by Lee \etal
+\cite{Lee09}). Indoor Manhattan environments are a sub--class of
+general Manhattan environments; examples are shown in figure
+\figref{indoor-manhattan}. This is an attractive hypothesis class
+because
 \begin{enumerate}
-\item{it captures many \textit{regularities} within man--made
-  environments;}
+\item{it captures many regularities within man--made environments;}
 \item{the geometric primitives (floor/wall/ceiling) are immediately
@@
 
+\begin{figure}[tb]
+  \centering
+  \subfloat[]{
+    \includegraphics[width=0.3\textwidth]{figures/example-model1}
+  }
+  \subfloat[]{
+    \includegraphics[width=0.3\textwidth]{figures/example-model2}
+  }
+  \subfloat[]{
+    \includegraphics[width=0.3\textwidth]{figures/example-floorplan}
+  }
+  \caption{(a-b) Examples of indoor Manhattan environments. The red
+    line illustrates the seam representation. (c)
+    An indoor Manhattan environment viewed from above.}
+  \label{fig:indoor-manhattan}
+\end{figure}
+
 An indoor Manhattan model is much like an archicetural floor--plan,
-and can be specificied as a set of 2D line segments representing walls
-together with the position of the floor and ceiling plane (figure
-XXX). In this paper we adopt an image--domain parametrisation due to
-its convenience for inference and learning. Following Flint \etal
-\cite{Flint11} we represent hypotheses as a path running from the left
-edge to the right edge of the image, which we will refer to as a
-\textit{seam}. A seam $\Seam$ consists of a sequence of pairs of
-scalars,
+and can be specified as a set of 2D line segments representing walls
+together with the position of the floor and ceiling plane
+(\figref{indoor-manhattan}). In this paper we adopt an image--domain
+parametrisation due to its convenience for inference and
+learning. Following Flint \etal \cite{Flint11} we represent hypotheses
+as a path running from the left edge to the right edge of the image,
+which we will refer to as a \textit{seam}. A seam $\Seam$ consists of
+a sequence of pairs of scalars,
 \begin{equation}
@@
 \end{equation}
-where $\seam_i$ is the $y$--cordinate at which the path intersects
-image column $i$, $\Orient_i$ is the orientation of the wall in that
+where $\seam_i$ is the $y$--coordinate at which the path intersects
+image column $i$, $\Label_i$ is the orientation of the wall in that
 column, and $\Width$ is the image width. Remarkably, this simple
@@
 of indoor Manhattan models, there are two properties of the seam
-representation that will be of relevance to the following
+representation that will be relevant in the following
 sections. Firstly, images of indoor Manhattan environments can be
 rectified so that vertical lines in the world project to vertical
-lines in the image \cite{Flint10eccv}. Secondly, indoor Manhattan
-models decompose in the following manner. We said earlier that a seam
-$\Seam$ specifies a unique 3D reconstruction. Let
-$\Orient(x,y;\Seam)\in\TheOrientations$ be the orientation of the 3D
-surface projecting to pixel $(x,y)$. Then Flint \etal
-\cite{Flint10eccv} showed that $\Orient$ is functionally dependant
-\textit{only} on the labels for column $x$. That is, we may write
+lines in the image \cite{Flint10eccv}. Secondly, the mapping from seam
+to 3D reconstruction decomposes in the following manner. We said
+earlier that a seam $\Seam$ specifies a unique 3D reconstruction. Let
+$\Orient(x,y;\Seam)$ be the orientation of the 3D
+surface projecting to pixel $(x,y)$. Flint \etal \cite{Flint10eccv}
+showed that $\Orient$ is functionally dependant \textit{only} on the
+pair $(\seam_x,\Label_x)$ for column $x$. That is, we may write
 \begin{equation}
@@
 \end{equation}
-Similarly, letting $\Depth$ be the depth of the reconstruction at
-$(x,y)$ we may write $\Depth$ in terms of a single label pair,
-\cite{Flint10iccv}
+Similarly, letting $\Depth(x,y,;\Seam)$ be the distance from the
+camera to the surface projecting to $(x,y)$ we may write
 \begin{equation}
@@
 
-\subsection{Observation Model}
+\subsection{Feature Space}
 
 We adopt the probabilistic model formulated by Flint \etal
-\cite{Flint11}, which relates indoor Manhattan hypotheses to
-single--view image features, multiple--view photoconsistency terms,
+\cite{Flint11}, which relates indoor Manhattan reconstructions to
+single--view image features, multiple--view photo--consistency terms,
 and a reconstructed point cloud. From our perspective these are simply
-features and any subset may be omitted if in appropriate for a given
-application.
-
-TODO: walk through the graphical model?
-
-Under this model the prior on environments is a geometric distribution
-in the number of distinct wall segments, which means, roughly, that we
-consider reconstructions with more walls apriori less likely. The
-benefit of this prior is that its logarithm is linear in the number of
-wall segments,
+features and any subset may be omitted if inappropriate for a given
+application. Under this model the prior on reconstructions is
 \begin{equation}
-  \log P(\Seam) = - \vect{n} \cdot \Penalties
+  \log P(\Seam) = - \vect{n} \cdot \Penalties + O(1)
+  \label{eq:prior}
 \end{equation}
@@
 of various categories (concave, convex, occluding) and $\Penalties$ is
-a vector of hyper--parameters.
+a hyper--parameter. We may interpret \eqnref{prior} as,
+roughly, that reconstructions with more walls are less likely.
 
-Flint \etal showed that for a variety of single--view and
-multiple--view features $\Features$ there are reasonable choices of
-likelihood that can be written in the form
+Flint \etal \cite{Flint11} showed that for a variety of features
+$\Feature$ (including single--view and multiple--view) there are
+reasonable choices of likelihood that can be written
 \begin{equation}
-  \log P(\Features ~|~ \Seam) = \sum_{x=1}^\Width 
-  \PixelPayoff_{\Features}(x,\seam_x)
+  \log P(\Feature ~|~ \Seam) = \sum_{x=1}^\Width 
+  \PixelPayoff_{\Feature}(x,\seam_x)
+  \label{eq:likelihood}
 \end{equation}
-where $\PixelPayoff_{\Features}$ is a 2--dimensional matrix computed
-deterministically from $\Features$ \cite{Flint11}. It is easy to show
-that for each likelihood in \cite{Flint11}, the log--likelihoods
-is linear in the hyper--parameters $\LikelihoodHypers$,
+where $\PixelPayoff_{\Feature}$ is a real matrix computed
+deterministically from $\Feature$ and $\PixelPayoff_{\Feature}(i,j)$
+is the element at row $i$, column $j$. It is easy to show that for each
+sensor model described in \cite{Flint11} the log--likelihood is linear
+in the hyper--parameters $\LikelihoodHypers$, \ie
 \begin{equation}
-  \log P(\Features ~|~ \Seam) = 
+  \log P(\Feature ~|~ \Seam) = 
   \sum_{x=1}^\Width
-    \LikelihoodHypers_{\Features} 
+  \LikelihoodHypers_{\Feature} 
   \cdot
-    \PixelFtr_{\Features}(x,\seam_x)
+  \PixelFtr_{\Feature}(x,\seam_x) ~.
 \end{equation}
-
-Assuming conditional independence between the feature types, we have
-for the posterior,
+Assuming conditional independence between features, the posterior is
 \begin{eqnarray}
-  P(\Seam ~|~ \Features_1 \ldots \Features_n)
+    P(\Seam ~|~ \Features)
   &\propto&
-  P(\Seam) P(\Feature_1 ~|~ \Seam) \ldots P(\Feature_n ~|~ \Seam)\\
-  \log P(\Seam ~|~ \Features_1 \ldots \Features_n)
+    P(\Seam) \prod_{i=1}^n P(\Feature_i ~|~ \Seam)\\
+    \log P(\Seam ~|~ \Features)
   &=&
@@
     + \sum_{i=1}^n \sum_{x=1}^\Width 
-      \LikelihoodHypers_i
-      \cdot
-      \PixelFtr_i(x,\seam_x)
+    \LikelihoodHypers_i \cdot \PixelFtr_i(x,\seam_x)
     + O(1) ~.
@@
 
-In order to place our model within a structured prediction framework
-we need to re--write the above in terms of a joint feature space
-$\JointFtr(\Seam,\Features)$. The decomposability of indoor Manhattan
-mdoels into payoff matrices permits precisely such a
+So far, this model closely follows that described in \cite{Flint11};
+our contribution is to place this model into a statistically rigorous
+learning framework. To do this we need to rewrite the above in terms
+of a joint feature function $\JointFtr(\Features,\Seam)$ and a
+parameter vector $\Model$. The decomposability of indoor Manhattan
+models into payoff matrices permits precisely such a
 formulation. Defining
 \begin{align}
-  \JointFtr(\Seam,\Features) =
+  \JointFtr(\Features,\Seam) =
   \begin{bmatrix}
@@
 \end{align}
-we see that \eqnref{log-posterior} can be written
+we see that \eqnref{log-posterior} can be written as
 \begin{equation}
-  \log P(\Seam ~|~ \Features_1 \ldots \Features_n)
+  \log P(\Seam ~|~ \Features)
   =
-  \bigl\langle \Model ~,~ \JointFtr(\Seam,\Features) \bigr\rangle + O(1)
+  \bigl\langle \Model ~,~ \JointFtr(\Features,\Seam) \bigr\rangle +
+  O(1)
+  \label{eq:log-linear-posterior}
 \end{equation}
@@
 $\langle\cdot,\cdot\rangle$ denoting an inner product. Since $\Model$
-contains all the free parameters in the model, the goal learning will
-be to optimise $\Model$ with respect to a set of training features
-$\{\Features_i\}$ together with ground truth labellings $\{\Seam_i\}$.
+contains all free parameters in the model, the goal of learning will
+be to optimise $\Model$ with respect to a training set
+$\{(\Features_i,\Seam_i)\}$.
 
@@
 
-The precise make--up of the feature space depends on the particular
-forms for the likelihoods and which sensor modalities are
-available. We define separate feature spaces for the single-- and
-multiple--view contexts; both are summarized in \figref{featurespace}.
+The precise make--up of the feature space depends on the available
+sensor modalities. We define separate feature spaces for the single--
+and multiple--view contexts; these are summarized in
+\figref{featurespace}.
 
@@
   \centering
-  \begin{tabular}{@{}llll@{}}
+  \begin{tabular}{@{}lllll@{}}
     \toprule
-    Feature & Dimensionality & Single--view? & Reference \\
+    Feature & Dimensionality & Multi--view? & Single--view? & Reference \\
     \midrule
-    Line sweeps & 1 & yes & Lee \etal \cite{Lee09}\\
-    Stereo photo--consistency & 4 & no & Flint \etal \cite{Flint11}\\
-    Point cloud & 2 & no & Flint \etal \cite{Flint11}\\
+    Stereo photo--consistency & 4 & yes & no & Flint \etal \cite{Flint11}\\
+    Point cloud & 2 & yes & no & Flint \etal \cite{Flint11}\\
+    Line sweeps & 1 & yes & yes & Lee \etal \cite{Lee09}\\
+    RGB+HSV & 6 & no & yes & \\
+    Gabor responses\footnotemark & 12 & no & yes & \\
     \bottomrule
   \end{tabular}
-  \caption{The composition of our feature space. Each row corresponds
-    to one or more of the $\PixelFtr_i$ in \eqnref{joint-ftr-def}.}
+  \caption{The composition of our single-- and multiple--view feature
+    space. We omit color and Gabor features from the multiple--view
+    feature space for training efficiency.}
   \label{fig:featurespace}
@@
 
-A loss function $\Delta(\Seam,\EstSeam)$ is a measure of the
-deviation of a predicted label $\EstSeam$ from the ground truth
-$\Seam$. In the context of learning one often faces a trade--off
-between choosing a loss that leads to tractible optimization, and
-choosing a loss that measures what one ``really'' cares about. For
-example, Hoiem \etal \cite{Hoiem05} implicitly use a per--pixel loss,
-since this decomposes optimization pixel--wise. However, it would be
-hard to argue that pixel--wise accuracy is an honest measure of what
-the authors care about: some pixel--level errors are insignificant to
-the overall quality of a reconstruction, while others are
-catastrophic. Whether the authors care about the accuracy of their
-reconstruction in 3D, its utility for contextual object detection, or
-even the aesthetic appeal of reconstructions, a per--pixel loss is
-almost certainly a weak measure of success.
+\footnotetext{4 orientations, 3 scales}
 
-A more reasonable choice would be the relative depth error, which has
-been the gold standard within the reconstruction community for decades
-\cite{Hartley04}, and measures the average deviation between
-reconstructed and ground truth depths:
+Next we define a loss function $\Delta(\Seam,\EstSeam)$ measuring the
+cost of predicting some reconstruction $\EstSeam$ when in fact the
+true reconstruction is $\Seam$. In the context of learning one often
+faces a trade--off between choosing a loss that leads to tractable
+optimization, and choosing a loss that measures the quantity that one
+``really'' cares about. For example, Hoiem \etal \cite{Hoiem05} learn
+a per--segment orientation classifier, then pass this as input to a
+separate 3D reconstruction system \cite{Hoiem2005}. However, what one
+``really'' cares about is some loss defined on the output of the
+entire system rather than the output of individual components, since
+some segment--level mistakes are insignificant to the overall
+reconstruction quality, while others are catastrophic. This is not a
+criticism of the authors' choice, but an illustration of the
+trade--off faced when choosing a loss. In this paper we show how to
+learn efficiently with respect to a loss defined on the final
+reconstruction.
+
+The relative depth error has been the gold standard within the
+reconstruction community for more than a decade \cite{Hartley04}, and
+measures the average deviation between reconstructed and ground truth
+depths. In our notation,
 \begin{equation}
+  \label{eq:depth-loss}
   \DepthLoss(\Seam,\EstSeam)
@@
   \frac
-        {\Depth(\Pixel;\Seam) - \Depth(\Pixel;\EstSeam)}
+      {\abs{\Depth(\Pixel;\EstSeam) - \Depth(\Pixel;\Seam)}}
       {\Depth(\Pixel;\Seam)} ~,
@@
 \begin{equation}
-  \LblLoss(\Seam,\EstSeam)
+  \label{eq:label-loss}
+  \LblLoss(\EstSeam,\Seam)
   =
@@
   \sum_{\Pixel}
-    \bigl[~\Orient(\Pixel;\Seam) \neq \Orient(\Pixel;\EstSeam)~\bigr] ~,
+  \bigl[~\Orient(\Pixel;\EstSeam) \neq \Orient(\Pixel;\Seam)~\bigr] ~,
 \end{equation}
 where $[p]$ is 1 if $p$ is true and 0 otherwise. An attractive
-characteristic of the indoor Manhattan class is that both of these
-loss functions can be optimized exactly. The algorithmic details for
-this are left to \sectref{learning}; the key result that we need to
-attend to here is that $\DepthLoss$ and $\LblLoss$ can be written in a
-form similar to the payoff formulation for the feature
-likelihoods. Recall the functional independence established in
-\eqnref{func-indep}.
+characteristic of the indoor Manhattan class is that \textit{both of
+  these losses can be optimized exactly}. The algorithmic details are
+left to \sectref{learning}; the key result we establish here is that
+$\DepthLoss$ and $\LblLoss$ can be written in a form resembling the
+payoff formulation \eqnref{likelihood} for the feature likelihoods.
 
-In the
-case of reconstruction, the gold standard is the relative depth error
-\cite{Hartley04}
+First we invoke the independence established in \eqnref{depth-indep}:
+\begin{equation}
+  \DepthLoss(\EstSeam,\Seam)
+  =
+  \frac{1}{N}
+  \sum_{x=1}^{\Width}
+  \sum_{y=1}^{\Height}
+  \frac
+      {\abs{\tilde{\Depth}(x,y;\hat{\seam}_x) - \Depth(x,y;\Seam)}}
+      {\Depth(x,y;\Seam)} ~.
+\end{equation}
+Defining a real matrix $\LossMat_{\Seam}$,
+\begin{equation}
+  \LossMat_{\Seam}(x,j) =
+  \sum_{y=1}^H
+  \frac
+      {\abs{\tilde{\Depth}(x,y;j) - \Depth(x,y;\Seam)}}
+      {\Depth(x,y;\Seam)} ~,
+\end{equation}
+we see that we can write $\DepthLoss$ in the form
+\begin{equation}
+  \DepthLoss(\EstSeam,\Seam)
+  =
+  \frac{1}{N}
+  \sum_{x=1}^{\Width} \LossMat_{\Seam}(x,\hat{\seam}_x) ~.
+  \label{eq:depth-loss-matrix}
+\end{equation}
+There is a similar form for $\LblLoss$, which we omit here due to
+space constraints.
 
-Should we show how to write these as payoff matrices here, or leave
-this to the learning section?
+\subsubsection{Choosing a Loss Function}
 
-L1, L2 error: easy representation, not meaningful, give examples
-
-Labelling error
--> showing these will require some of the ICCV framework to have been presented
--> give nice intuition, show scatter plots, examples
-
-Relative depth error
--> showing these will require some of the ICCV framework to have been presented
--> give nice intuition, show scatter plots, examples
-
-None of these are pixel classifiers!
-
-Writing loss functions in the seam representation -> maybe just
-mention that this will be useful later
-
-\section{Inference}
-\label{sec:inference}
-
-Describe the inference algorithm in terms of payoff/penalty
-optimization, cite ECCV -> this section must be short!
+Neither of the above losses is unequivocally the ``correct'' loss; the
+choice will depend on the application. One might expect a strong
+correlation between the losses, and indeed one can show analytically
+that
+\begin{equation}
+  \DepthLoss(\EstSeam,\Seam) = 0
+  \iff
+  \LblLoss(\EstSeam,\Seam) = 0 ~.
+\end{equation}
+However, in our experiments we found only a weak correlation between
+these losses away from the origin. For example, the scatter plot shown
+in figure \figref{comparative-scatter} shows a significant number of
+outliers that score very well on $\DepthLoss$ but poorly on
+$\LblLoss$, and vice versa.
 
@@
 
-Risk minimization formulation
-%  -> min_f E[ Loss(f(x),y*) ]
-Empirical risk problem
-%  -> min_f sum_{xi} Loss(f(x),y*)
-Write f as a linear argmax
-%  -> f(x) = argmax_y <w,psi(x,y)>
-Reduced to quadratic optimization (cite SVMstruct papers)
-%  -> min_f ...
--> subject to ....
-Minimize using algorithm in ...
-This requires a separation oracle, which is the topic of the
-following subsection
+We turn now to the problem of learning within the model described
+above. Our learning task is to identify a prediction function
+$\Predictor$ mapping observed features $\Features$ to reconstructions
+$\Seam$. We seek the loss minimizer
+\begin{equation}
+  \OptPredictor = 
+  \argmin_{\Predictor} 
+  \Expected \Bigl[ 
+    \Loss\bigl(\Predictor(\Features),\Seam\bigr)
+    \Bigr] ~,
+\end{equation}
+which we approximate in the framework of empirical risk minimization as
+\begin{equation}
+  \OptPredictor = 
+  \argmin_{\Predictor} 
+  \sum_k \Loss\bigl(\Predictor(\Features_k),\Seam_k\bigr) ~,
+\end{equation}
+where $k$ indexes a training set. To perform this optimization we
+turn to the tools of structured prediction \cite{Bakir07}, and in
+particular the structured SVM \cite{Tsochantaridis04}. First we need
+to define $\Predictor$. In this paper we consider predictors of the
+form
+\begin{equation}
+  \Predictor_{\Model}(\Features) = \argmax_{\Seam} 
+  \bigl\langle \Model ~,~ \JointFtr(\Features,\Seam) \bigr\rangle
+  ~.
+  \label{eq:predictor-class}
+\end{equation}
+Comparing with \eqnref{log-linear-posterior} we see that each
+predictor of the form \eqnref{predictor-class} is simply implementing
+MAP inference under some set of hyper--parameters $\Model$. We now
+turn to the optimisation problem itself. Following the standard
+approach \cite{Bakir07} we cast the learning problem as a constrained
+optimisation problem,
+\begin{equation}
+  \begin{split}
+    \min_{\Model,\xi} &
+      \hspace{2mm} 
+    \frac{1}{2} \|\Model\|^2 +
+      C \sum_{k=1}^n \xi_k\\
+    s.t. & \hspace{2mm} \forall k, \Seam \neq \Seam_k:~
+      \bigl\langle\Model, \JointFtr(\Features_k,\Seam_k)\bigr\rangle -
+      \bigl\langle\Model, \JointFtr(\Features_k,\Seam)\bigr\rangle
+      \geq
+      \Loss(\Seam,\Seam_k) - \xi_k ~.
+  \end{split}
+  \label{eq:svm-problem}
+\end{equation}
+Tsochantaridis \etal \cite{Tsochantaridis04} described an algorithm
+for solving this minimisation that is now used extensively within
+machine learning and computer vision. To apply this algorithm here we
+must solve two inference problems:
+\begin{enumerate}
+  \item{\textit{Prediction.} This is the maximization described in
+    \eqnref{predictor-class}.}
+  \item{\textit{Separation.} The algorithm described in
+    \cite{Tsochantaridis04} requires a user--supplied procedure to
+    find the ``most--violated constraint'' at each iteration. That is,
+    \begin{equation}
+      \argmax_{\Seam}
+        \bigl\langle\Model, \JointFtr(\Features_k,\Seam)\bigr\rangle
+        + \Loss(\Seam,\Seam_k) ~.
+        \label{eq:separation-problem}
+    \end{equation}
+  }
+\end{enumerate}
 
-Loss-augmented inference
--> the problem
--> writing loss the functions as payoff matrices
--> writing a payoff matrix for loss-augmented inference
--> write a proof that ``the optimal model payoff matrix
-B=... solves the loss-augmented inference problem <w,psi> + L
+Our solutions to both of the above build on the algorithm presented by Flint
+\etal \cite{Flint10eccv,Flint11}, which is a dynamic programming
+solution to problems of the form
+\begin{equation}
+  \argmax_{\Seam}
+  \sum_x \PixelPayoff(x,\seam_x) - \sum_j \CornerPenalty(j;\Seam) ~.
+  \label{eq:inference-problem}
+\end{equation}
 
-\section{Practical Considerations}
-\label{sec:practical-considerations}
+\subsection{Inference (Prediction)}
 
-- negative slacks are more harmful than one would expect
+We showed in \secref{model} that \eqnref{inference-problem} can be
+written in the form \eqnref{predictor-class}, so the prediction
+problem is a straightforward application of \cite{Flint10eccv}. This
+is as expected since, as we have already remarked,
+\eqnref{predictor-class} is equivalent to MAP inference on indoor
+Manhattan reconstructions, which was precisely the subject
+of \cite{Flint10eccv}.
 
-- inconsistent inference is very harmful
+\subsection{Loss--Augmented Inference (Separation)}
 
-- inference when the ground truth not in the hypothesis class
+It turns out that the separation problem can also be solved using the
+dynamic programming algorithm mentioned above, as the following
+proposition shows.
 
-- conditioning in a joint feature space
+\begin{proposition}
+  Let $(\Features_k,\Seam_k)$ be a training instance with payoff
+  matrices $\{\PixelPayoff_i\}$ as defined in \eqnref{likelihood}.
+  Let
+  \begin{equation}
+    \AugPayoff = \LossMat_{\Seam_k} + \sum_i \PixelPayoff_i ~.
+    \label{eq:aug-payoff}
+  \end{equation}
+  Then the solution to \eqnref{inference-problem} with
+  $\PixelPayoff=\AugPayoff$ is identical to the solution to
+  \eqnref{separation-problem}.
+\end{proposition}
+\begin{proof}
+  Direct equivalence of the expressions to be maximized. First
+  substitute \eqnref{log-linear-posterior} and
+  \eqnref{depth-loss-matrix} into \eqnref{inference-problem}:
+  \begin{equation}
+    \log P(\Seam ~|~ \Features) +
+    \sum_{x=1}^{\Width} \LossMat_{\Seam_k}(x,\seam_x) ~
+  \end{equation}
+  Further substituting \eqnref{likelihood} and defining
+  $\CornerPenalty$ as in \cite{Flint11} gives
+  \begin{equation}
+    \sum_i \sum_{x=1}^{\Width} \PixelPayoff_i(x,\seam_x) - 
+    \sum_j \CornerPenalty(j;\Seam) +
+    \sum_{x=1}^{\Width} \LossMat_{\Seam_k}(x,\seam_x) ~.
+  \end{equation}
+  Finally we see that substituting \eqnref{aug-payoff} gives
+  \begin{equation}
+    \sum_{x=1}^{\Width} \AugPayoff(x,\seam_x) - 
+    \sum_j \CornerPenalty(j;\Seam)
+  \end{equation}  
+\end{proof}
 
@@
 
+We evaluated our approach on the data--set proposed in \cite{Flint11},
+which consists of 18 sequences of six environments averaging 59
+seconds in duration. We sampled key--frames at regular intervals. Each
+``instance'' in our training and hold--out sets consists of one base
+frame together with four auxiliary frames.
+
+We compared with the bootstrapping approach described in
+\cite{Flint11}. Our metrics differ from theirs in two ways. Firstly,
+they compute relative depth error using the maximum of the ground
+truth and estimated depths in the denominator, whereas we always use
+the ground truth in the denominator. These metrics are separated by at
+most a monotonic transform but the latter is more convenient to
+represent in our framework. Secondly, when we compute labelling error
+we differentiate vertical and horizontal surfaces only, whereas they
+also differentiate the two vertical orientations. The latter approach
+makes a side--by--side comparison difficult because the two vertical
+orientations are symmetric and their labels can always be
+interchanged.
+
+The performance of these two algorithms are summarized in
+\figref{mv-performance}. Our approach significantly out--performs the
+bootstrapping algorithm. Anecdotally we noticed that much of the
+improvement resulted from a reduction in catastrophic failures. This
+makes sense because we would expect the learning algorithm to
+concentrate on reducing those mistakes that result in the largest
+loss. Some example predictions are shown in \figref{results-pics};
+many more are included in additional material.
+
+\begin{figure}[tb]
+  \centering
+  \begin{tabular}{@{}p{20mm}p{20mm}p{20mm}p{20mm}p{20mm}@{}}
+    \toprule
+     & \multicolumn{2}{c}{Depth Error (\%)}
+     & \multicolumn{2}{c}{Labelling Error (\%)} \\
+    \midrule
+    Sequence & This Paper\footnotemark[2] & Flint \etal 
+             & This Paper\footnotemark[3] & Flint \etal \\
+    \midrule
+    \tt{ground}   & 4.9    & 66.6    & 2.9    & 10.4  \\
+    \tt{foyer1}   & 6.1    & 6.6     & 3.1    & 3.1   \\
+    \tt{foyer2}   & 4.3    & 5.4     & 3.7    & 4.0   \\
+    \tt{corridor} & 14.6   & 52.9    & 9.5    & 19.2  \\
+    \tt{mcr}      & 34.0   & 67.6    & 15.    & 16.2  \\
+    \tt{kitchen}  & 16.8   & 23.6    & 5.2    & 6.1   \\
+    \midrule
+    Average       & \textbf{13.4} & 37.1 & \textbf{6.7} & 9.8   \\
+    \bottomrule
+  \end{tabular}
+  \caption{Multiple--view reconstruction performance on held--out
+    data, comparing with Flint \etal \cite{Flint11}. For unavoidable
+    reasons we use slightly different metrics so our figures differ
+    from those published in \cite{Flint11}. See main text for
+    explanation.}
+  \label{fig:mv-performance}
+\end{figure}
+
+\begin{figure}[tb]
+  \centering
+  \begin{tabular}{@{}p{20mm}p{20mm}p{20mm}p{20mm}p{20mm}@{}}
+    \toprule
+     & \multicolumn{2}{c}{Depth Error (\%)}
+     & \multicolumn{2}{c}{Labelling Error (\%)} \\
+    \midrule
+    Sequence & This Paper\footnotemark[2] & Flint \etal 
+             & This Paper\footnotemark[3] & Flint \etal \\
+    \midrule
+    \tt{ground}   & 17.3   & 24.5    & 7.8    & 12.2  \\
+    \tt{foyer1}   & 25.1   & 31.0    & 15.1   & 22.2  \\
+    \tt{foyer2}   & 29.1   & 30.1    & 15.9   & 18.6  \\
+    \tt{corridor} & 31.7   & 33.6    & 19.3   & 24.8  \\
+    \tt{mcr}      & 70.1   & 45.9    & 26.7   & 20.8  \\
+    \tt{kitchen}  & 25.1   & 26.2    & 7.7    & 11.9  \\
+    \midrule
+    Average       & 33.1   & \textbf{31.9} & \textbf{15.4} & 18.4   \\
+    \bottomrule
+  \end{tabular}
+  \caption{Single--view reconstruction performance on held--out data,
+    comparing with Flint \etal \cite{Flint10eccv}.}
+  \label{fig:sv-performance}
+\end{figure}
+
+\begin{figure}[tb]%
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/comparative_scatter}
+  \caption{The effect of the loss function on training. We train two
+    predictors, one with respect to $\DepthLoss$ and one with respect
+    to $\LblLoss$, then evaluate both on all held--out instances. Each
+    data point shows the error obtained by one predictor on held--out
+    instance. The differing distribution of errors shows that the two
+    predictors trade off errors as expected.}
+  \label{fig:comparative-scatter}
+\end{figure}
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/psi_evolution}
+  \caption{Evolution of $\Model$ during training. Each series shows
+    the value of one component of $\Model$. After an
+    exploration phase the model converges.}
+  \label{fig:psi-evolution}
+\end{figure}
+
 \section{Single View Results}
 \label{sec:sv-results}
+
+We evaluated our system for single--view reconstruction using the same
+data--set described in the previous section. We used the single--view
+features summarized in \figref{featurespace}. We compared our approach
+to the single--view approach of Flint \etal \cite{Flint10eccv}, which
+uses the same dynamic programming algorithm that we rely upon, but
+uses hand--tuned features.
+
+Performance for each algorithm is summarized in
+\figref{mv-performance}. When measured by labelling error, our
+approach out--performs the hand--tuned weights, but on the depth error
+metric our approach is inferior. While investigating this result we
+found that our learning algorithm assigns small weights to all but the
+line--sweep features, which are the same features used by
+\cite{Flint10eccv}. This suggests that the hand--tuned weights are in
+fact close to optimal within this feature space, though one would
+expect that with additional feature engineering our learning algorithm
+would be able to leverage further salient information and reduce the
+error rate.
+
+\footnotetext[2]{This column represents the predictor trained with
+  respect to $\DepthLoss$.}
+
+\footnotetext[3]{This column represents the predictor trained with
+  respect to $\LblLoss$.}
+
+\section{Discussion}
+\label{sec:discussion}
+
+The hypothesis class considered in this paper is amongst the most
+complex (in terms of internal constraints on the output space) studied
+within the structured prediction framework. In this section we turn to
+some practical lessons learnt that may be of value to other
+practitioners.
+
+\subsubsection{Condition in the joint feature space, not the input
+  feature space.}
+
+A common pre--processing operation for statistical learning is to
+transform the observed features $\Features$ to zero mean and unit
+variance. However, for structured prediction tasks it is the joint
+feature space $\JointFtr$ that should be conditioned:
+\begin{equation}
+  \label{eq:conditioning}
+  \JointFtr' = \frac{\JointFtr - \mu}{\sigma^2} ~.
+\end{equation}
+Ideally one would sample from the joint feature space to determine the
+conditioning transformation, but the distribution of inputs and
+outputs is generally unknown in an empirical risk minimization
+setting. Instead, we use the training set as a proxy. We compute the
+empirical mean and variance of
+$\{\JointFtr(\Features_k,\Seam_k)\}_{k=1}^n$ at the outset, then apply
+the transformation \eqnref{conditioning} after each feature
+computation.
+
+\subsubsection{Condition the loss terms.}
+
+For any $\eta>0$, the minimization problem \eqnref{svm-problem} is
+equivalent (under the substitution $\Model'=\eta\Model$,
+$\xi'=\eta\xi$) to:
+\begin{equation}
+  \begin{split}
+    \min_{\Model',\xi'} &
+      \hspace{2mm} 
+    \frac{1}{2} \|\Model'\|^2 +
+      \eta C \sum_{k=1}^n \xi_k'\\
+    s.t. & \hspace{2mm} \forall k, \Seam \neq \Seam_k:~
+      \bigl\langle\Model', \JointFtr(\Features_k,\Seam_k)\bigr\rangle -
+      \bigl\langle\Model', \JointFtr(\Features_k,\Seam)\bigr\rangle
+      \geq
+      \frac{1}{\eta} \Loss(\Seam,\Seam_k) - \xi_k' ~.
+  \end{split}
+  \label{eq:equiv-problem}
+\end{equation}
+Although any $\eta>0$ preserves the correctness of the optimization
+algorithm, we found that choosing $\eta=\mbox{Var}(\Loss)$ improved
+numerical stability, since this means the loss terms will have roughly
+unit variance. Unfortunately, we cannot use the training set to
+estimate $\mbox{Var}(\Loss)$ since the loss for the ground truth
+reconstruction is always zero. Instead we computed
+$\Loss(\Features_k,\Seam_j)$ for each $k \neq j$ in the training
+set. This is not an ideal estimate, but we found that it worked well
+in practice.
+
+\subsubsection{Check that the hypothesis class contains the ground
+  truth.}
+
+The algorithm described in \cite{Tsochantaridis04} implicitly assumes
+that the hypothesis class $\mathcal{Y}$ contains the ground truth
+labels $\Seam_k$. This means that if $\AugSeam$ is the maximizer of
+\eqnref{separation-problem} then
+\begin{equation}
+  \bigl\langle\Model, \JointFtr(\Features_k,\Seam_k)\bigr\rangle
+  - \bigl\langle\Model, \JointFtr(\Features_k,\AugSeam)\bigr\rangle
+  - \Loss(\AugSeam,\Seam_k) 
+  \leq 
+  0 ~,
+  \label{eq:positive-slacks}
+\end{equation}
+since otherwise we would have
+\begin{equation}
+  \bigl\langle\Model, \JointFtr(\Features_k,\Seam_k)\bigr\rangle
+  + \Loss(\Seam_k,\Seam_k) 
+  >
+  \bigl\langle\Model, \JointFtr(\Features_k,\AugSeam)\bigr\rangle
+  + \Loss(\AugSeam,\Seam_k) ~,
+\end{equation}
+contradicting $\AugSeam$ as the maximizer of
+\eqnref{separation-problem}. However, our output space contains
+fundamentally real--valued quantities such as polygon vertices, which
+are recovered only to some finite precision by the inference
+algorithm, and since our ground truth labels were acquired by manual
+labelling, they sometimes exceed the maximum precision of the
+inference algorithm. In this case we effectively have $\Seam_k \notin
+\mathcal{Y}$ (although there is always some $\Seam' \in \mathcal{Y}$
+close to $\Seam_k$), so it is possible that $\AugSeam$ violates
+\eqnref{positive-slacks}. Our workaround here is simply to check the
+condition \eqnref{positive-slacks} each time we solve the separation
+problem and, if violated, substitute $\Seam_k$ for $\AugSeam$. This is
+justified by the observation that if \eqnref{positive-slacks} is
+violated for $\AugSeam$ then it is violated for all
+$\Seam\in\mathcal{Y}$. One could think of this as learning with
+respect to the hypothesis class $\mathcal{Y} \union \{\Seam_k\}$ but
+evaluating with respect to $\mathcal{Y}$. This is not an ideal
+solution but we found it to work well in practice. Unfortunately this
+patch has the side--effect of hiding bugs in the inference algorithm,
+so care is warranted.
 
@@
 
-Wow, we really did rock didn't we?
+We have presented a unified learning framework for reconstructing
+polygonal models from single and multiple views of a scene. We have
+chosen to work with the indoor Manhattan class of models in order to
+leverage the parametrisation and inference algorithm recently proposed
+for this hypothesis class \cite{Flint10eccv,Flint11}. Our approach to
+learning performs a single optimisation with respect to a clearly
+defined loss function. Experiments show our system out--performing the
+state--of--the--art for multiple--view reconstruction (by a large
+margin) and on one metric for single--view reconstruction.
 
-Hypothesis class, why it's interesting
+In future work will extend this approach to learn geometry
+together with scene classifiers and context--aware object detectors,
+optimising with respect to a single joint loss function.
 
-Statistical rigour is guuuud
+\newcommand{\CompFrame}[3]{\includegraphics[width=0.18\textwidth]{figures/comparison_frames/#1_frame#2_#3.png}}
 
-These fields are ready for this kind of unification
+\begin{figure}[tb]%
+  \centering
+  \begin{tabular}{cccccc}
+    \begin{sideways}
+      $~~~\DepthLoss$
+    \end{sideways} &
+    \CompFrame{lab_foyer1}{002}{depthtrained} &
+    \CompFrame{lab_ground1}{012}{depthtrained} &
+    \CompFrame{exeter_mcr1}{012}{depthtrained} &
+    \CompFrame{exeter_mcr1}{042}{depthtrained} &
+    \CompFrame{som_corr1}{022}{depthtrained} 
+    \\
 
-Rigorous joint learning for contextual object detection: a challenge
+    \begin{sideways}
+      $~~~\LblLoss$
+    \end{sideways} &
+    \CompFrame{lab_foyer1}{002}{lbltrained} &
+    \CompFrame{lab_ground1}{012}{lbltrained} &
+    \CompFrame{exeter_mcr1}{012}{lbltrained} &
+    \CompFrame{exeter_mcr1}{042}{lbltrained} &
+    \CompFrame{som_corr1}{022}{lbltrained} 
+    \\
+
+    \begin{sideways}
+      Flint \etal
+    \end{sideways} &
+    \CompFrame{lab_foyer1}{002}{iccv} &
+    \CompFrame{lab_ground1}{012}{iccv} &
+    \CompFrame{exeter_mcr1}{012}{iccv} &
+    \CompFrame{exeter_mcr1}{042}{iccv} &
+    \CompFrame{som_corr1}{022}{iccv} 
+    \\
+
+    \begin{sideways}
+      $~~$Ground
+    \end{sideways}
+    \begin{sideways}
+      $~~~$Truth
+    \end{sideways} &
+    \CompFrame{lab_foyer1}{002}{gt} &
+    \CompFrame{lab_ground1}{012}{gt} &
+    \CompFrame{exeter_mcr1}{012}{gt} &
+    \CompFrame{exeter_mcr1}{042}{gt} &
+    \CompFrame{som_corr1}{022}{gt} 
+  \end{tabular}
+  \caption{Multiple--view reconstructions predicted by our system
+    (held--out samples). The first two rows represent the
+    predictors trained on $\DepthLoss$ and $\LblLoss$ respectively,
+    the third row is from \cite{Flint11}, and the fourth row is ground
+    truth.}
+  \label{fig:results-pics}
+\end{figure}
 
 \bibliographystyle{splncs}
-\bibliography{AVLstrings,VisionRefs}
+\bibliography{AVLstrings,VisionRefs,MendeleyRefs}

### 2012-03-12 papers/eccv2012/main.tex
ECCV additional material submitted

@@
   feature space.}
-
+gi
 A common pre--processing operation for statistical learning is to
@@
       \geq
-      \frac{1}{\eta} \Loss(\Seam,\Seam_k) - \xi_k' ~.
+      %%% TODO: in the line below 1/eta should be replaced with eta --
+      %%% this was incorrect in the review submission!!
+      \eta \Loss(\Seam,\Seam_k) - \xi_k' ~.
   \end{split}

### 2012-06-01 papers/eccv2010/rebuttal/rebuttal-v1.txt

We thank all three reviewers for their positive comments as well as constructive criticism.

A key concern raised by all three reviewers is the apparently restrictive nature of the proposed indoor Manhattan models. Since this constitutes by far the most frequent objection, and indeed is the only substantial complaint of Reviewer 5, we dedicate the following four paragraphs to responding to this issue.

It must be stressed that our aim is to recover _coarse_, _high-level_ models in order to obtain useful geometric cues that will assist further reasoning, rather than to recover intricately detailed 3D models of the environment. Given the ill-posed nature of the single-view reconstruction problem, it seems obvious to us that there is a trade-off between the complexity of models on one hand and the robustness of the estimation process on the other. We believe that indoor Manhattan model represent an ideal point on this trade-off curve since they capture an excellent set of geometric properties (such as the dominant surface orientations, height of the floor and ceiling w.r.t the camera, and the position of major room boundaries) while retaining a low-dimensional parameterization.

In addition to the robustness argument above, there is a further advantage to our use of indoor Manhattan models: the position of room boundaries, which our models recover explicitly, will prove more salient for purposes such as navigation, scene classification, or contextual priming of object detection, than the detailed geometry of room clutter. On this view the choice of indoor Manhattan models constitutes an effort to recover room boundaries _in spite_ of scene clutter,  and indeed our system is able to achieve precisely this (see rows 1-7 of Figure 2 in supplementary material).

We further note that previous attempts at single-view reconstruction have (necesarily) encountered the same robustness/complexity trade-off, and most have indeed chosen to restrict the class of scene models. Lee [4] and Flint [5] employ exactly the same class of models as us, while Hedau et al. [12] choose the far more restricted class of simple cuboids. Hoiem et al. [10] also restricts surface orientations to three dominant directions; only Saxena et al. [11] allow more general models, though they do not evaluate their system on difficult indoor scenes.

It seems appropriate to briefly sketch some concrete examples of "high-level reasoning systems" that would benefit from being able to recover indoor Manhattan models. A few examples are:
- Contextual cues for object detection, in the style pioneered by Torralba ("Contextual priming for object detection", IJCV 2003) and demonstrated for outdoor scenes by Hoiem ("Putting objects in perspective", IJCV, 2008). It seems obvious that "distance from the ground", "heigh relative to floor-ceiling distance", and "distance from nearest wall" would make extremely powerful cues when detecting objects in indoor scenes.
- Categorizing scenes into semantic categories such as "kitchen" or "bedroom", such as that investigated by Fei-Fei et al. ("A bayesian hierarchical model for learning natural scene categories", CVPR 2005). Incorporating cues such as the dimensions and shape of a room, as well as taking surface boundaries and orientation into account when reasoning from photometric cues, would constitute a powerful addition to texture-only approaches taken previously.
- High-level descriptions of room boundaries would find many uses for the purpose of path-planning and navigation in mobile robotics.

Reviewer 2 requests that we include the optimal substructure proof in the main text. We agree and have re-worked the layout of the final paper in order to achieve this.

We also thank reviewer 2 for pointing out the work of Barinova et al. Their work does not overlap with our approach (they use CRFs together with an EM algorithm) nor is a comparison with their results appropriate (their approach is suited only to outdoor scenes); we do, however, regret that this citation was omitted from the review manuscript and have added it to the final version. We also thank reviewer 3 for the reference to the work of Felzenszwalb and Vexler, but since this work is to appear at CVPR later this year we obviously could not have cited it in the review manuscript since it was not available at that time. Furthermore, our approach is quite different to theirs.

In response to the question raised by reviewer 2 regarding the ligitimacy of our evaluation framework, we respond that we chose the pixelwise evaluation metric in line with that chosen by Hoiem [10], Lee [4], and Hedau [12], and in order to allow direct comparisons with those approaches.

Reviewer 5 construes our approach as a "smoother" of the provided orientation estimate. Perhaps the reviewer has misunderstood the relationship of our work to that of Lee et al: We employ their plane sweep heuristic to obtain noisy orientation estimates as depicted in the second column of figures 3, 4, 5, and 6 (supplementary material). Importantly, our system does _not_ use the output of the branch-and-bound algorithm they employ; rather, our dynamic programming solution represents an alternative to their algorithm and we _compare_ our results to theirs.

Finally, to answer reviewer 5's question about our ground truth acquisition: it was obtained by hand-labelling structure-from-motion maps as described in the first paragraph of section 7. Small inaccuracies may arise from human error or structure-from-motion errors, but we see no evidence for systematic bias here.

### 2012-06-01 papers/eccv2010/rebuttal/rebuttal.txt

We thank the reviewers for positive comments and constructive criticism.

A key concern raised by the reviewers is the apparently restrictive nature of the indoor Manhattan model. We therefore stress that our proposed method recovers the dominant 3D structure of the environment while remaining robust to clutter. The gain in complexity compared to the state-of-the-art (Lee et al. [4]) is substantial, dropping from exponential to linear. Our approach also provides high-level geometric features, which are not simply pixel-wise orientations but meaningful large planar regions in the environment, encoding strong constraints for many tasks such as scene classification and contextual priming of object detection.

Our choice of the indoor Manhattan model constitutes an effort to recover room boundaries _in spite_ of scene clutter, and indeed our system is able to achieve precisely this (see rows 1-7 of Fig 2 in supp. material). We hope these figures explicitly answer the criticisms of R2 and R5 regarding robustness to clutter.

Previous attempts at single-view reconstruction have (necessarily) encountered the same robustness/complexity trade-off, and most have indeed chosen to restrict the class of scene models. [4] and [5] employ exactly the same class of models as us, while [12] choose the far more restricted class of simple cuboids. [10] also restrict surface orientations to three dominant directions; only [11] allow more general models, though they do not evaluate their system on indoor scenes.

We regret overlooking the work of Barinova et al pointed out by R2, and will rectify this in the final version. We thank R3
for highlighting the latest work by Felzenszwalb and Veksler that will appear at CVPR2010.

R2 and R5 question our evaluation framework. While there are indeed issues with the pixel-wise accuracy metric, it was previously used by Hoiem et al. [10], Lee et al. [4], and Hedau et al. [12], and so we adopted it to allow direct comparisons with those approaches. To answer the question raised by R5 regarding the method of ground truth acquisition: we used hand-labelled structure-from-motion maps as ground truth (lines 293-296). There may be small human- or structure-from-motion errors, but we do not believe there is systematic bias. R3 noted that our framework "very convincingly shows superiority over Lee et al" and was impressed by our automated ground truth acquisition mechanism.

R2 noted that the optimal substructure proof was interesting and should be included in the main paper. We will include the proof as recommended.

R5 construes our approach as a "smoother" of the initial orientation estimate. To clarify the relationship of our work to Lee et al: we employ their plane sweep heuristic to obtain noisy orientation estimates as depicted in the second column of Figs 3,4,5,6 (supp. material), but our system does _not_ use the output of the branch-and-bound algorithm they employ; instead, our dynamic programming solution represents a much more efficient alternative to their algorithm and we _compare_ our results to theirs.

### 2012-06-01 papers/eccv2010/main.tex

\newcommand\Model{\mathcal{M}}
\newcommand\eqnref[1]{\eqref{eq:#1}}
\newcommand\Hcf{H_{c \rightarrow f}}
\newcommand\Hcfhat{\hat{H}_{c \rightarrow f}}

\begin{abstract}
A number of recent papers have investigated the Manhattan world
assumption, in which surfaces in the world are assumed to be aligned
with one of three dominant directions
\cite{Coughlan99,Zhang02,Furukawa09,Lee09}. In this paper we present a
dynamic programming solution to the reconstruction problem for
``indoor'' Manhattan worlds (a sub--class of Manhattan worlds). Our
algorithm deterministically finds the global optimum and exhibits
computational complexity linear in both model complexity and image
size. This is an important improvement over previous methods that were
either approximate \cite{Furukawa09} or exponential in model
complexity \cite{Lee09}. We present results for a new dataset
containing several thousand manually annotated images, which are
released in conjunction with this paper.
\end{abstract}

\section{Introduction}

In this paper we investigate the problem of reconstructing geometric
models from single images. We choose to fit a model containing sparse
geometric primitives with the intended aim of higher--level
reasoning. If novel--view synthesis or photo-realistic 3D modelling
were the goal then a dense reconstruction may be appropriate, but if
the model is to be used as input to some reasoning problem, as is the
motivation for the present work, then a simpler model may be more
useful. In particular, when reasoning about indoor environments the
location of the floor, ceiling, and wall planes provide strong cues to
the probable locations of objects and events, whereas a raw point
cloud would be more difficult to interpret. Furthermore, simple models
have fewer degrees of freedom and are easier to infer from ambiguous
image evidence.

It is perhaps for this reason that the past few years have seen
considerable interest in the Manhattan world assumption
\cite{Coughlan99,Zhang02,Lee09,Furukawa09,Flint10}, in which each
surface is assumed to have one of three possible orientations. Making
this assumption introduces regularities that can improve the quality
of the final reconstruction \cite{Furukawa09}. Several papers have
investigated the even more constrained class of \textit{indoor}
Manhattan scenes \cite{Lee09,Flint10,Hedau09}, which consist entirely of
vertical walls extending between the floor and ceiling planes. A
surprisingly broad set of interesting environments can be modelled
exactly or approximately as indoor Manhattan scenes \cite{Flint10}. It
is with this class of scenes that this paper is concerned.

The present work describes a novel and highly efficient algorithm to
obtain models of indoor Manhattan scenes from single images using
dynamic programming. In contrast to point cloud reconstructions, our
algorithm assigns semantic labels such as ``floor'', ``wall'', or
``ceiling''. We show that our method produces superior results when
compared to previous approaches. Furthermore, our algorithm exhibits
running time linear in both image size and model complexity (number of
corners), whereas all previous methods that we are aware of
\cite{Lee09,Flint10} are exponential in model complexity.

The remainder of the paper is organised as follows. Section 2
describes previous work in this area and section 3 outlines our
approach. In section 4 we pose the indoor Manhattan problem formally,
then in section 5 we develop the dynamic programming solution. We
present experimental results in section 6, including a comparison with
previous methods. Concluding remarks are given in the final section.

\section{Background}

Many researchers have investigated the problem of recovering
polyhedral models from line drawings. Huffman \cite{Huffman71}
proposed a scheme to distinguish concave, convex, and occluding lines,
which allows discrimination between possible and impossible
objects. Waltz \cite{Waltz72} investigated a more general problem
involving incomplete line drawings and spurious measurements from
shadows. Sugihara \cite{Sugihara82} proposed an algebraic approach to
interpreting line drawings, while the ``origami world'' of Kanade
\cite{Kanade80} utilised heuristics to reconstruct a model of hollow
shells and planar sheets. These approaches were limited to synthetic
line drawings.

Hoiem \etal \cite{Hoiem05} and Saxena \etal \cite{Saxena09} have
investigated the single image reconstruction problem from a machine
learning perspective. Their approaches assign pixel--wise orientation
labels using appearance characteristics of outdoor scenes. Hedau \etal
\cite{Hedau09} extend this to indoor scenes, though their work is
limited to rectangular box environments.

The work most closely related to our own is that of Lee \etal
\cite{Lee09}, who have shown that line segments can be combined to
generate indoor Manhattan models using a branch--and--bound
algorithm. In contrast to their approach, our system uses dynamic
programming to efficiently search \textit{all} feasible indoor
Manhattan models rather than just those generated by line segments. As
a result we obtain more accurate models, can reconstruct more complex
environments (see \figref{positive-egs}), and obtain computation times
several orders of magnitude less than their approach, as will be
detailed in \sectref{results}.

Furukawa \etal \cite{Furukawa09} have used the Manhattan world
assumption for stereo reconstruction. Their models are of high
quality but their quoted computation times (of between one minute and
one hour forty minutes) make their approach unsuitable to on--line
applications. Their approach is not comparable to ours because they
use multiple calibrated views, and they search a different class of
models.

\section{Outline of Proposed Approach}

The Manhattan world assumption states that world surfaces are oriented
in one of three mutually orthogonal directions \cite{Coughlan99} and
the indoor Manhattan assumption further states that the environment
consists of a floor plane, a ceiling plane, and a set of walls
extending vertically between them \cite{Lee09}. Each wall therefore
has one of two possible orientations, and each corner is either
concave, convex, or occluding (see \figref{fin-egs}). Indoor Manhattan
models are interesting because they can represent many indoor
environments approximately or exactly, yet they introduce regularities
to the reconstruction problem that makes possible a left--to--right
decomposition of the scene, on which the dynamic programming algorithm
developed in this paper rests. Our approach to reconstructing indoor
Manhattan environments consists of the following five steps:
\begin{enumerate}
  \item{Identify dominant directions. (\sectref{vpts})}
  \item{Identify the floor and ceiling planes. (\sectref{fcmap})}
  \item{Warp the image to a canonical view. (\sectref{warp})}
  \item{Obtain weak orientation estimates. (\sectref{orient-est})}
  \item{Estimate the final model. (Sections 4 and 5)}
\end{enumerate}

\subsection{Identifying dominant directions}
\label{sect:vpts}

We identify three dominant directions by estimating mutually
orthogonal vanishing points in the image. We follow a similar approach
to Koseck\`{a} and Zhang \cite{Zhang02}, in which $k$--means
clustering is used to obtain an initial estimate, after which EM is
employed to identify the final vanishing points
$\vect{v_1}$,$\vect{v_3}$, and $\vect{v_3}$.

If intrinsic camera parameters are not provided then we construct the
camera matrix $K$ from the detected vanishing points by assuming that
the camera centre is at the image centre and choosing a focal length
and aspect ratio such that the calibrated vanishing points are
mutually orthogonal. For the remainder of this paper
$\vect{v_1}$,$\vect{v_2}$, and $\vect{v_3}$ denote the calibrated
vanishing points.

We assume that the vanishing point with largest absolute
$y$--coordinate represents the up--down direction in the world (\ie
the direction in which gravity operates), and by convention we always
label this direction $\vect{v_3}$. The horizon line is given by
$\vect{h}=\vect{v_1}\times\vect{v_2}$.

\subsection{Identifying the floor and ceiling planes.}
\label{sect:fcmap}

An indoor Manhattan scene has exactly one floor and one ceiling plane,
both with normal direction $\vect{v_3}$. It will be useful in the
following sections to have available the mapping $\Hcf$ between the
image locations of ceiling points and the image locations of the floor
points that are vertically below them (see \figref{fcmap}). $\Hcf$ is
a planar homology with axis $\vect{h}$ and vertex $\vect{v_3}$
\cite{Criminisi01} and can be recovered given the image location of any
pair of corresponding floor/ceiling points $(\vect{x_f},\vect{x_c})$
as
\begin{equation}
  \Hcf = I + \mu\frac{\vect{v_3}\vect{h}^T}{\vect{v_3} \cdot \vect{h}} ~,
\end{equation}
where $\mu = <\vect{v_3},\vect{x_c},\vect{x_f},
\vect{x_c}\times\vect{x_f}\times\vect{h}>$ is the characteristic cross
ratio of $\Hcf$.

\begin{figure}[tb]
\centering
\includegraphics[width=0.4\textwidth]{figures/fcmap}
\caption{The mapping $\Hcf$ transfers points between the ceiling and
  floor.}
\label{fig:fcmap}
\end{figure}

Although we do not have \textit{a priori} any such pair
$(\vect{x_f},\vect{x_c})$, we can recover $\Hcf$ using the following
RANSAC algorithm. First, we sample one point $\vect{\hat{x}_c}$ from
the region above the horizon in the Canny edge map, then we sample a
second point $\vect{\hat{x}_f}$ collinear with the first and
$\vect{v_3}$ from the region below the horizon. We compute the
hypothesis map $\Hcfhat$ as described above, which we then score by
the number of edge pixels that $\Hcfhat$ maps onto other edge pixels
(according to the Canny edge map). After repeating this for a fixed
number of iterations we return the hypothesis with greatest score.

Many images contain either no view of the floor or no view of the
ceiling. In such cases $\Hcf$ is unimportant since there are no
corresponding points in the image. If the best $\Hcf$ output from the
RANSAC process has a score below a threshold $k_t$ then we set $\mu$
to a large value that will transfer all pixels outside the image
bounds. $\Hcf$ will then have no impact on the estimated model.

\subsection{Warping the image to a canonical view}
\label{sect:warp}

The algorithms presented in the remainder of this paper will be
simplified if vertical lines in the world appear vertical
in the image. We therefore warp images according to a homography
\begin{equation}
  H = 
  \begin{pmatrix}
    \vect{v_3} \times \vect{e_3} \\
    \vect{v_3} \\
    \vect{v_3} \times \vect{e_3} \times \vect{v_3}
  \end{pmatrix}
  ~,
\end{equation}
where $\vect{e_3}=[0,0,1]^T$.

This step is not crucial to our approach and we could instead do all
geometric calculations in the original image. Hence, the fact that the
warped image will be contain a singularity if $\vect{v_3}$ is within
the image bounds does not restrict the applicability of our
approach. For the sake of simplicity, however, we present the
remainder of this paper assuming that $\vect{v_3}$ is outside the
image bounds.

\subsection{Obtaining weak orientation estimates}
\label{sect:orient-est}

Our algorithm requires a pixel--wise surface orientation estimate to
bootstrap the search. Obtaining such estimates has been explored by
several authors \cite{Lee09,Hoiem05,Saxena09}. We adopt the simple and
efficient line--sweep approach of Lee \etal \cite{Lee09}, which
assigns one of four labels $\{1,2,3,z\}$ to each pixel, where $z$
represents an unknown orientation and the remaining labels identify
the normal direction of the surface that generated the pixel. We
denote the orientation map $o:\mathbb{R}^2\rightarrow \{1,2,3,z\}$.

Note that our algorithm is is not dependent on the manner in which $o$
is obtained; any method capable of estimating surface orientations
from a single image, including the work of Hoiem \cite{Hoiem05} or
Saxena \cite{Saxena09}, could be used instead.

We generate four binary images $B_i$ such that pixel $B_i(\vect{x})=1$
if and only if $o(\vect{x})=i$. We then compute the integral image
\cite{Viola01} for each $B_i$, which allows us to count the number of
pixels of a given orientation within any rectangular sub--image in
$O(1)$ time. This represnetation expediates evaluation of the cost
function described in the following section.

\section{Problem statement}
\label{sect:probstmt}

We represent an indoor Manhattan model $\Model$ as an alternating
sequence of corners and walls $\{c_1,W_1,c_2,W_2,...,W_{n-1},c_n\}$,
$c_i < c_{i+1}$. Each corner $c_i$ is the column index at which two
walls meet, and each wall $W_i=(r_i,a_i)$ comprises an orientation
$a_i\in\{1,2\}$, which determines whether its vanishing point is
$\vect{v_1}$ or $\vect{v_2}$, and a row index $r_i$ at which its upper
edge meets the corner to its left. An example model is shown in
\figref{model-params}. Remarkably, this simple description completely
specifies all four corners of each wall segment (see
\figref{pqrs}). Clockwise from top--left the vertices of the i\th wall
are
\begin{eqnarray}
  \vect{p_i} &=& [c_i,r_i,1]^T \\
  \vect{q_i} &=& \vect{p_i} \times \vect{v_{a_i}} \times
       [1,0,-c_{i+1}]^T \\
  \vect{r_i} &=& H_{c \rightarrow f} \vect{q_i} \\
  \vect{s_i} &=& H_{c \rightarrow f} \vect{p_i} ~.
\end{eqnarray}

\begin{figure}[tb]%
  \centering
  \subfloat[][]{%
    \label{fig:pqrs}
    \includegraphics[width=0.6\textwidth]{figures/pqrs}}
  \subfloat[][]{%
    \label{fig:model-params}
    \includegraphics[width=0.4\textwidth]{figures/model-params}}
  \caption{(a) An illustration of the model
    $\Model=\{c_1,(r_1,a_1),...,(r_4,a_4),c_5\}$; (b) the values
    $(c_i,r_i,a_i,c_{i+1})$ fully specify the four corners of a wall.}
\end{figure}

A model $\Model$ predicts one of three possible surface orientations
$\mathcal{L}=\{1,2,3\}$ for each pixel $\vect{x}$. We denote this
prediction $\Model(\vect{x})\in\mathcal{L}$ where
\begin{equation}
  \Model(\vect{x}) =
  \begin{cases}
    a_i & \mbox{if } \vect{x} \in
    \mbox{quad}(\vect{p_i},\vect{q_i},\vect{r_i},\vect{s_i}) \\
    a_v & \mbox{otherwise}
  \end{cases}
  ~,
\end{equation}
where $a_v$ is the index of the vanishing point representing the
vertical direction in the world.

Not all models $\Model$ are physically realisable, but those that are
not are easily discarded using simple tests on the locations of walls
and vanishing points as enumerated by Lee \etal \cite{Lee09}. The
reader is referred to their paper for details; we simply note that a
model is feasible if all of its corners are feasible, and the
feasibility of a corner is dependent only on the immediately adjoining
walls.

We are now ready to state the minimisation problem formally. Given an
input image of size $W \times H$ and an orientation map $o$, we define
the cost of assigning pixel $\vect{x}$ label $l \in \mathcal{L}$ as
\begin{equation}
  C_p(\vect{x},l) =
  \begin{cases}
    0, & \mbox{if } o(\vect{x}) = l \mbox{ or } o(\vect{x}) = z \\
    1, & \mbox{otherwise}\\
  \end{cases}
  ~.
\end{equation}
The overall cost for a model $\Model$ containing $n$ corners is then
\begin{equation}
  \label{eq:cost}
  C(\Model) = \lambda n - \sum_{\vect{x}\in\mathcal{I}} C_p(\vect{x},\Model(\vect{x}))
\end{equation}
where $n\lambda$ is the regularisation term. We seek the model
$\Model^*$ with minimum cost
\begin{equation}
  \label{eq:opt-model}
  \Model^* = \underset{\Model}{\operatorname{argmax}} ~ C(\Model)
  ~.
\end{equation}

Finally, the marginal cost $C_w$ of the $i$\th wall $W_i=(r_i,a_i)$ is
\begin{equation}
  C_w = \sum_{x_i\leq x\leq x_{i+1}} \Bigl(
  \sum_{1 \leq y \leq U_i} C_p(x,y,a) +
  \sum_{U < y \leq L_i} C_p(x,y,a_v) +
  \sum_{L < y \leq H} C_p(x,y,a) \Bigr)
\end{equation}
where $L_i$ and $U_i$ are respectively the ceiling/wall and floor/wall
intersection at column $x_i$. The three sums over $y$ are $O(1)$
operations given the integral images described in
\sectref{orient-est}.

\section{Proposed algorithm}
In this section we present a dynamic programming solution to the
problem posed in \sectref{probstmt}. We develop the algorithm
incrementally to assist readability. Our solution rests on a
decomposition of the indoor Manhattan problem into a set of
sub--problems that can be solved in terms of one another. Consider the
following sub--problem.

Let $f_{in}(x,y,a,k)$ be the cost of a model $\Model^+ =
\{c_1,W_1,...,W_{n-1},c_n\}$ such that
\begin{enumerate}
  \item{$n=k$, (\ie $\Model^+$ contains $k$ corners),}
  \item{$c_n=x$ (\ie the rightmost corner is at column $x$),}
  \item{$W_{n-1}=(y,a)$
    (\ie wall $n-1$ terminates at row $y$ with orientation
    $a$),}
  \item{$\Model^+$ is feasible, and}
  \item{$\Model^+$ has minimal cost among all such models.}
\end{enumerate}

We show in the additional material that if a model
\begin{equation}
  \Model=\{c_1,W_1,...,W_{k-1},c_k\}
\end{equation}
is a solution to the sub--problem $f_{in}(c_k,r_k,a_k,k)$, then the the
truncated model
\begin{equation}
  \Model'=\{c_1,W_1,...,W_{k-2},c_{k-1}\}
\end{equation}
is a solution to the sub--problem $f_{in}(c_{k-1}, r_{k-1}, a_{k-1},
k-1)$.

In light of this fact we introduce the following recurrence relation:
\begin{equation}
  \label{eq:fin-recurrence-first}
  f_{in}(x,y,a,k) = \min_{x'<x,~y',a'} \Bigl( f_{in}(x',y',a',k) +
  C_w \Bigr) ~,
\end{equation}
where $C_w$ is the marginal cost of the wall $W=(y',a')$. We also
introduce the following boundary conditions:
\begin{equation}
  \label{eq:fin-base}
  f_{in}(x,y,a,k) =
  \begin{cases}
    0, & \mbox{if } x=0 \\
    \infty, & \mbox{if } k<0 \mbox{ and } x>0
  \end{cases}
  ~,
\end{equation}
which are justified, respectively, because a model with rightmost
corner at $x=0$ does not span any part of the image, so its cost must
be zero, and because a model cannot contain less than zero corners.

Finally, the cost of the optimal model \eqnref{opt-model} is
\begin{equation}
  \label{eq:opt-entry-point}
  C(\Model^*) = 
  \min_{{1\leq y \leq H \atop k \leq K} \atop a\in\{1,2\}}
  \Bigl( f_{in}(W,y,a,k) - \lambda k \Bigr) ~.
\end{equation}
where $K$ is a parameter specifying the maximum model complexity and
$\lambda$ is the per--wall penalty.

We compute $C(\Model^*)$ by recursively evaluating $f_{in}$
according to \eqnref{fin-recurrence-first} until we reach one of the
boundary conditions \eqnref{fin-base}. In line with the standard
dynamic programming approach we cache each evaluation to avoid
evaluating the same node multiple times. Once all nodes have been
evaluated, $\Model^*$ is obtained by back--tracking through the
evaluation graph.

\begin{figure}[tb]
\centering
\includegraphics[width=0.75\textwidth]{figures/fin-hypotheses}
\caption{Three models that satisfy constraints 1--4 for the
  sub--problem $f_{in}(x,y,a,k)$. There will only be one such model
  that satisfies the minimal cost constraint. Note that the
  right--most wall in each model terminates at $(x,y)$ with
  orientation $a$.}
\label{fig:fin-egs}
\end{figure}

\textbf{Complexity.} Since each node is evaluated at most once, the
overall complexity is given by the product of the number of nodes and
the complexity of evaluating each node. The latter is $O(W^2H)$, since
the minimisation in \eqnref{fin-recurrence-first} is over $O(WH)$
terms and computing the marginal cost $C_w$ requires $O(W)$
additions. The number of nodes is $O(WHK)$, so the overall complexity
of the basic algorithm is $O(W^3H^2K) = O(L^5K)$ where $L=\max(W,H)$.

\subsection{Auxiliary nodes}

The basic algorithm is hampered by the need to minimise simultaneously
over $x'$ and $y'$ in \eqnref{fin-recurrence-first}. In this section
we show how to decouple these minimisations by introducing auxiliary
sub--problems, which results in a significant reduction in
computational complexity.

We introduce three new sub--problems: $f_{up}$, $f_{down}$, and
$f_{out}$. Each of these is defined identically to $f_{in}$, except
that constraint 3 is replaced according to the following table:

\begin{tabular}{p{0.1\textwidth}p{0.75\textwidth}}
  $f_{up}$ & $r_{n-1} \leq y$ (\ie the rightmost wall terminates above
  row $y$) \\
  $f_{down}$ & $r_{n-1} \geq y$ (\ie the rightmost wall terminates 
  below row  $y$) \\
  $f_{out}$ & appending a wall $W=\{y,a\}$ to $M^+$ would result in a
  feasible model. \\
\end{tabular}

Using a similar argument to that given in the previous section it can
be shown that these sub--problems are related by the following
recurrence relations:
\begin{eqnarray}
  \label{eq:fout-recurrence}
    f_{out}(x,y,a,k) &=& \min_{a'\in\{1,2\}} \min
      \begin{cases}
        f_{up}(x,y,a',k) & \\
        f_{out}(x,y,a',k) & \\
        f_{down}(x,y,a',k) &
      \end{cases}
\end{eqnarray}
\begin{eqnarray}
  \label{eq:fup-recurrence}
  f_{up}(x,y,a,k) &=& 
  \begin{cases}
    \min \Bigl(f_{in}(\cdot), f_{up}(x,y-1,a,k)\Bigr), &
    \mbox{if } y \geq 1 \\
    \infty, & \mbox{otherwise}
  \end{cases}
\end{eqnarray}
\begin{eqnarray}
  \label{eq:fdown-recurrence}
  f_{down}(x,y,a,k) &=& 
  \begin{cases}
    \min \Bigl(f_{in}(\cdot), f_{down}(x,y+1,a,k)\Bigr), &
    \mbox{if } y \leq H \\
    \infty, & \mbox{otherwise}
  \end{cases}
\end{eqnarray}
\begin{eqnarray}
  \label{eq:fin-recurrence-basic}
  f_{in}(x,y,a,k) &=&
  \begin{cases}
    0, & \mbox{if } x=0 \\
    \infty, & \mbox{if } k<0 \\
    \min_{x'<x}\Bigl(f_{out}(x',v(x'),a,k-1)+c(x')\Bigr), &
    \mbox{otherwise}
  \end{cases}
\end{eqnarray}
where in the final equation $v(x')$ is the $y$--coordinate at which
the line through $\vect{v_a}$ and $(x,y)$ meets column
$x'$. Feasibility is enforced by removing either or both of the $f_{up}$
or $f_{down}$ terms in \eqnref{fout-recurrence} if such a corner would
result in an infeasible model.

\subsection{The line--jump optimisation}
Evaluating a $f_{in}$ node remains an $O(W)$ operation due to the
minimisation over $x'$ in \eqnref{fin-recurrence-basic}. In this
section we show how to reduce this to an $O(1)$ operation.

\begin{figure}[tb]
\centering
\includegraphics[width=0.4\textwidth]{figures/pixel-residuals}
\caption{A line from $\vect{x}$ to $\vect{v_a}$, and the distances $d$
  to nearby pixel centres (green dots). The starred pixel is the first
  that satisfies $d<\alpha$.}
\label{fig:pixel-residuals}
\end{figure}

Consider \figref{pixel-residuals}, showing a line from a pixel at
$\vect{x}$ to a vanishing point. If the line passed exactly through
some other pixel centre $\vect{x_p}$ then we could simply evaluate
$f_{in}(x_p,y_p,a,k)$ and terminate the line search at that point. It
is unlikely that the line will pass exactly through an integer--valued
pixel centre, so we instead specify a threshold $\alpha$ within which
a pixel centre will be considered to fall exactly on a line. We suffer
no loss of precision by doing this since the image itself is precise
only up to pixelation. We now have
\begin{equation}
  \label{eq:fin-recurrence}
  f_{in}(x,y,a,k) = \min
  \begin{cases}
    \min_{x_p \leq x' < x}\Bigl(f_{out}(x',v(x'),a,k-1)+c(x')\Bigr) & \\
    f_{in}(x_p,v(x_p),a,k-1)+c(x_p) & \\
  \end{cases}
  ~.
\end{equation}
where $(x_p,y_p)$ is the closest pixel to $(x,y)$ satisfying $x_p<x$
and for which $d=\vect{x} \cdot (\vect{x_p} \times \vect{v_a})$ is
less than $\alpha$. Both $(x,y)$ and $\vect{v_a}$ have integer
coordinates since they are observed at pixel locations, so we have the
upper bound
\begin{equation}
  x-x_p<\beta
\end{equation}
for some constant $\beta$, which is dependent only on our choice for
$\alpha$ \cite{Schrijver98}.

\textbf{Complexity.} Evaluating each node is now an $O(1)$
operation, so the overall complexity of our algorithm is given by the
number of nodes in the evaluation graph, which is
\begin{equation}
  O(KL^2) ~.
\end{equation}

\subsection{Down--sampling the orientation map}

As a further optimisation we down--sample the orientation map $o$,
replacing each $m \times m$ block of cells with a single cell
containing the majority vote over the block. This reduces both the
number of nodes in the graph and the cost of evaluating $C_w$.

\section{Other approaches}

\textbf{Graph cuts. }
\label{sect:comp-graph-cuts}
Many pixel--labelling problems have been successfully solved using
graph cuts. Kolmogorov and Zabih \cite{Kolmogorov02} have investigated
the class of cost functions that can be minimised using graph cuts.
Their work demonstrates that \textit{regularity}, a special case of
sub--modularity, is a necessary condition for any energy function to be
minimised via graph cuts.

It turns out that the cost \eqnref{cost} is not regular, so cannot be
minimised using graph cuts. Intuitively this is because the indoor
Manhattan constraint induces complicated dependencies between the
pixels in each column. \footnote{For example, if some pixel $\vect{p}$
  is assigned label $a$ then $\vect{q}=\Hcf\vect{p}$ must be assigned
  the same label, even though the two may be arbitrarily far from one
  another in the image.} Furthermore, applying graph cuts to this
problem would entail using a technique such as $\alpha$--expansion
\cite{Kolmogorov02}, which is both approximate and
non--deterministic. In contrast, our approach is exact and
deterministic.

\textbf{Branch and bound.}
Lee \etal \cite{Lee09} proposed a branch--and--bound solution to the
indoor Manhattan problem. Their approach identifies straight lines in
the image and then combines them in various permutations to generate a
set of model hypotheses, each of which are evaluated using a cost
function similar to \eqnref{cost}. Whereas their search space grows
combinatorially with the number of lines in the model, our approach is
linear in model complexity and hence we can model significantly more
complex scenes. We give timing comparisons with their approach in
\figref{time-vs-k}.

Their approach also differs from ours in that they only allow wall
boundaries to occur where lines are observed in the image. Our system
can easily be extended to enforce such a constraint by removing
$f_{in}$ terms from \eqnref{fup-recurrence} and
\eqnref{fdown-recurrence} for locations where lines are not
detected. However, we have found this to be unnecessary because the
cost \eqnref{cost} implicitly favours models with corners at the
location of observed lines since they will fit the orientation
estimate $o$ better. Furthermore, avoiding an explicit dependency on a
line detector is favourable because our system is not reliant on the
successful detection of structurally important line. In contrast, Lee
\etal are unable to build the correct model if many structurally
important edge are undetected. We have found that the structurally
important edges are often the least salient, since walls in a room are
often painted the same colour, so image gradients at their boundaries
are generated only by subtle lighting differences arising from the
different surface orientations.

\section{Results}
\label{sect:results}

We tested our system on a dataset of 634 manually annotated images of
indoor scenes. To expedite the annotation process we collected video
sequences and used structure--from--motion software to recover camera
poses, allowing us to project a manually specified floor plan into
each view.

In each experiment we computed the fraction of pixels for which the
orientation predicted by the output model $\Model$ agreed with the
ground truth orientation. Unless otherwise specified, the parameter
settings for the experiments below are $\alpha=0.01$, $K=7$, $m=4$,
$\lambda=100$.  Image sizes were 640 $\times$ 480 pixels.  We found
our approach to be very robust to all of these parameter values, as
the following experiments show.

We compared our results with the branch--and--bound approach of Lee
\etal \cite{Lee09}. In 138 of the images (21.7\% of the dataset),
their method was unable to estimate a building model as there was no
appropriate pair of line segments with which to initialise their
approach. In a pixel--wise evaluation their approach was able to
correctly label 54.3\% of pixels, while our approach obtained an
accuracy of 79.7\%. Omitting the images for which their approach was
unable to estimate a building structure, their approach obtained
68.1\% accuracy. We believe that the difficulty of our dataset (many
occluding objects, many images without a view of both floor and
ceiling) accounts for the significantly lower performance in
comparison to that quoted in \cite{Lee09}. Side--by--side comparisons
with their approach are included in additional material.

\subsection{Failure Cases}

\figref{negative-egs} shows four representative failure cases of our
approach. In the top--left panel the occlusion relationship between
two walls is incorrectly estimated, so the more distant wall is
thought to be occluding the closer wall. This is because the floor
patch in the bottom centre of the image is missed in the initial
orientation estimate. In the top--right panel, too few line segments
are detected and the initial orientation estimate is very poor. The
bottom--left panel shows an example of a chair that is wrongly
identified as part of a wall. The chair is aligned with the wall
behind it and this highlights the limitation of using only line
segments to estimate an initial orientation estimate. The
bottom--right panel shows how a deviation from the indoor Manhattan
assumption causes an incorrect model to be estimated. The exit sign
represents a vertical surface that does not extend from the the
ceiling to the floor, which our approach is currently unable to
handle.

\begin{figure}[tb]
\centering
\includegraphics[width=\textwidth]{figures/positive_montage}
\caption{Models estimated by our algorithm. Each panel contains three
  images: the original image, the initial orientation estimate, and
  the final model output by our system. Best viewed in colour.}
\label{fig:positive-egs}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=\textwidth]{figures/failure_montage}
\caption{Failure cases of our system. Best viewed in colour.}
\label{fig:negative-egs}
\end{figure}

\section{Conclusion}

We have shown that semantically meaningful models of indoor scenes can
be recovered efficiently for a range of Manhattan environments using
dynamic programming. Our approach is able to model complex scenes,
which would be intractable for previous methods that involved
combinatorial searches in the space of models. This work represents an
important increment on the state--of--the art both in terms of
accuracy and efficiency. Future work will investigate the use of
richer cues for obtaining initial orientation estimates since our
results indicate that is currently the limiting factor of our
approach.

\begin{figure}[tb]
\centering
\input{plots/time_vs_k}
\caption{Efficiency comparison with the branch--and--bound of Lee
  \etal \cite{Lee09} (our own implementation). Our approach was able to
  fit highly complex models in a under 2 seconds. Their algorithm took
  more than three minutes to fit a model of 3 corners; most of this
  time was spent evaluating the cost of the 200,000 building
  hypotheses.}
\label{fig:time-vs-k}
\end{figure}

\begin{figure}[tb]
\centering
\input{plots/jump_thresh}
\caption{The line--jump parameter $\alpha$ trades off accuracy
  (crosses) for efficiency (diamonds). Accuracy is computed relative
  to the baseline achieved when $\alpha=0$. The first few increments
  of $\alpha$ exhaust most of the efficiency gains, and at this end of
  the graph the loss in accuracy is negligible. Based on these results
  we set $\alpha=0.01$ in our remaining experiments.}
\label{fig:jump-thresh}
\end{figure}

\bibliographystyle{splncs}
\bibliography{AVLstrings,main}

### 2012-06-01 papers/iccv2011/main.tex

\begin{abstract}
  This paper addresses scene understanding in the context of a moving
  camera, integrating semantic reasoning ideas from monocular vision
  with 3D information available through structure--from--motion. We
  combine geometric and photometric cues in a Bayesian framework,
  building on recent successes leveraging the indoor Manhattan
  assumption in monocular vision. We focus on indoor environments and
  show how to extract key boundaries while ignoring clutter and
  decorations. To achieve this we present a graphical model that
  relates photometric cues learned from labeled data, stereo
  photo--consistency across multiple views, and depth cues derived
  from structure--from--motion point clouds. We show how to solve MAP
  inference using dynamic programming, allowing exact, global
  inference in $\sim$$100$ ms (in addition to feature computation of
  under one second) without using specialized hardware. Experiments
  show our system out--performing the state--of--the--art.
\end{abstract}

\section{Introduction}
Over the past decade, computer vision researchers working with
monocular images have pursued substantially different research
agendas to those working with multiple views. The focus for
monocular images has increasingly been to infer high--level facts
about the world, such as the locations of and interactions between
objects, semantic scene categories, and the spatial layout of the
environment. In contrast, much of the work concerning multiple views
has focused on reconstructing metric scene structure and camera poses
using techniques such as structure--from--motion, stereo, and
multiple--view stereo.

In this paper we leverage multiple view geometry for image
understanding purposes. We assume a moving camera with a
structure--from--motion system estimating its trajectory, and show how
to infer semantically meaningful models of the environment. We focus
on the \textit{indoor Manhattan representation}\cite{Lee09,FlintECCV10}, in which the world is
modeled in terms of floor, wall, and ceiling surfaces. This
representation captures many semantically meaningful aspects of the
environment, including (i) scale: the distance from floor to ceiling
suggests a scale for distances in the environment; (ii) boundaries:
walls constrain movement in the environment and suggest locations for
doors, windows, and other objects; (iii) gravity: the orientation of
the ground plane implies the direction in which gravity operates,
which constrains the arrangement of objects resting upon one another;
and (iv) shape: the organisation of walls in an environment suggests
a functional category (such as ``kitchen'' or ``office'').

\begin{figure}[tb]%
  \centering
  %\begin{tabular}{c}
    \includegraphics[width=0.2\textwidth]{figures/firstpage/lab_foyer2_frame041_dp.jpg}
    \includegraphics[width=0.2\textwidth]{figures/firstpage/exeter_mcr1_frame032_dp.jpg}
  %\end{tabular}
  \caption{Automatic reconstructions from our system.}
  %\label{fig:results-pics-top}
    \vspace{-5mm}
  \label{fig:firstpage}
\end{figure}

Indoor Manhattan models are useful because they capture these
properties explicitly, whereas to extract such properties from a dense
polygonal mesh would require additional non--trivial inference after
reconstruction. Our approach infers semantic properties of the scene
directly from multiple--view data, without an intermediate dense
reconstruction step. This makes sense if, as in our case, the semantic
properties constitute the ultimate goal of the system. Of course, if a
photo--realistic reconstruction is itself the end goal then our
approach is not suitable.

We build on recent work highlighting the efficacy of the indoor
Manhattan representation for single view reconstruction
\cite{Lee09,FlintECCV10}. Previous work employed a set of
heuristics as a cost function and was limited to monocular images. We
give a fully Bayesian account in which information from image
features, stereo, and 3D point clouds is integrated into a single MAP
optimization, and we learn all parameters from labeled examples. We
show that MAP inference in our model can be solved exactly and
efficiently ($\sim$$100$ ms per frame) using a generalization of the
dynamic programming algorithm of
\cite{FlintECCV10}.

%% Our work extends their approach to allow joint inference across
%% multiple views as well as to incorporate range data alongside
%% photometric cues. This involves a significant generalization of
%% their algorithm since neither of these new modalities are
%% representable using their pixel--wise cost function. We therefore
%% generalize their approach to a payoff--based formulation permitting
%% more general costs. We also show that in this new formulation one
%% of the state variables involved in their optimisation is
%% unecessary; its removal corresponds to eliminating an entire
%% dimension from the search and results in an order of magnitude
%% efficiency improvement.

%XXX Un-comment next para?
%When comparing our approach to previous work we propose that in
%addition to measuring geometric accuracy, the \textit{size} of
%alternative representations should also be compared, since models that
%capture key properties of the environment using fewer representational
%bits must have done a better job of filtering salient information
%from sensor data. This is in line with the relationship between the
%ability of an algorithm to compress data and its ability to make
%predictions (cite Hutter).

The remainder of this paper is organised as follows. Section two
places our contribution in context with related work. We introduce our
model in section three, then we describe inference in section four and
learning in section five. We give experimental results in section six,
followed by concluding remarks in section seven.

\begin{figure}[tb]%
  \centering
  \label{fig:ideal-models}
  \begin{tabular}{ccc}
    %\includegraphics[width=0.12\textwidth]{figures/true_models/lab_ground1_010_rect.png} &
    \includegraphics[width=0.12\textwidth]{figures/true_models/lab_foyer1_010_rect.png} &
    \includegraphics[width=0.12\textwidth]{figures/true_models/lab_kitchen_030_rect.png} &
    \includegraphics[width=0.12\textwidth]{figures/true_models/lab_ground1_030_rect.png} \\
    %\includegraphics[width=0.12\textwidth]{figures/true_models/lab_ground1_010} &
    \includegraphics[width=0.12\textwidth]{figures/true_models/lab_foyer1_010} &
    \includegraphics[width=0.12\textwidth]{figures/true_models/lab_kitchen_030} &
    \includegraphics[width=0.12\textwidth]{figures/true_models/lab_ground1_030}
  \end{tabular}
  \hspace{0.2cm}
  \caption{Three input images and the indoor Manhattan models we
    seek. Notice how each image column intersects exactly one wall.}
\end{figure}

\section{Background}
The Manhattan world assumption was introduced by Coughlan and
Yuille\cite{Coughlan99} over a decade ago and has seen increasing
attention in the computer vision literature over past years
\cite{Coughlan99,Zhang02,Lee09,Furukawa09,FlintECCV10}. Furukawa
\etal \cite{Furukawa09} proposed a Manhattan--world stereo algorithm
based on graph cuts. While their approach is concerned with dense
photo--realistic reconstructions, ours is intended to capture semantic
properties of the scene using a concise representation. The output of
their approach --- a polygonal mesh --- has no immediate semantic
interpretation, whereas our models, though less detailed, come
packaged with a direct interpretation. A by--product is efficiency: we
count computation time in hundreds of milliseconds, where as
Furukawa \etal report waiting more than an hour.

Another approach to interpreting Manhattan worlds is to model scenes
as a union of cuboids. This approach has a long history beginning with
Roberts' 1965 thesis \cite{Roberts65}, and has recently been revisited
using modern probabilistic techniques \cite{Gupta10,Vanegas10}.

Lee \etal \cite{Lee09} first proposed indoor Manhattan models (a
sub--class of general Manhattan models) for monocular
reconstructions. They used a branch--and--bound algorithm together
with a line--sweep heuristic for approximate inference. Flint \etal
\cite{FlintECCV10} employed a similar model but showed a
dynamic programming algorithm that performed exact inference in
polynomial time. In earlier work \cite{FlintCVPR10} Flint \etal also
demonstrated Manhattan reconstructions integrated with a SLAM system,
but this work inferred models from single frames and then extrapolated
these forward in time. In contrast, our work incorporates both
multiple view geometry and 3D points directly into a joint inference
procedure. We also learn parameters in a Bayesian framework, where as
neither Lee nor Flint utilized training data in any form.

Felzenszwalb and Veksler \cite{Felzenszwalb10} posed the
reconstruction problem in terms of energy minimization, which they
showed could be solved using dynamic programming, while Barinova \etal
\cite{Barinova08} modeled outdoor scenes using a CRF. However, these
approaches do not permit strong geometric constraints and so cannot be
extended to multiple views.

Semantic scene understanding has, broadly speaking, seen less
attention within the multiple view community. The CamVid
\cite{Brostow08} database of outdoor videos with semantic
segmentations is an important and encouraging exception. Brostow \etal
\cite{Brostow08} showed that simple structure--from--motion cues lead
to pleasing segmentations. Sturgess \etal \cite{Sturgess09} extended
this approach to a CRF framework. We compare our method with this
approach in \sectref{results}.

%XXX Do we need to cite work on object/action recognition, point cloud
%labelling?
% Object/action recognition
%   Leibe et al CVPR 2007 (best paper) SfM+object detection
%   Dala et al, ECCV 2006 HOG+flow
% Point labelling
%   Golovinskiy et al, ICCV 2009,  http://www.cs.princeton.edu/~funk/iccv09.pdf
%   Munoz et al, 2009 tech report: http://repository.cmu.edu/cgi/viewcontent.cgi?article=1329&context=robotics
% Our approach doesn't require 3D points at all, but can work with
% sparse OR dense points if available, and integrates with stereo +
% photometric features.

%XXX: The ECCV(?) paper on using manhattan + non-manhattan surfaces for
%reconstruction.

\section{Proposed Model}
\label{sect:model}

\begin{figure}[tb]%
  \centering
  \label{fig:corner-types}
  \includegraphics[width=0.3\textwidth]{figures/corner-types}
  \hspace{0.1cm}
  \caption{Each corner in an indoor Manhattan environment can be
    categorized as concave, convex, or occluding. Each vertical line
    intersects exactly one wall segment, the top and bottom of which
    we denote $\PfAtX$ and $\PcAtX$ respectively.}
\end{figure}

In this section we describe the indoor Manhattan model. We consider
three sensor modalities: monocular image features, stereo features,
and 3D point clouds. For each we present a generative model relating
observed features to the Manhattan scene structure, which we denote
$\Model$. For each sensor modality we show that MAP inference can be
reduced to maximization over a payoff function
$\PixelPayoff(x,y)$. This allows us to present a unified dynamic
programming solution in \sectref{inference}, which efficiently solves
MAP inference for all three sensor modalities.

%In this section we describe indoor Manhattan models and pose inference
%within this class as maximization of a payoff function that decomposes
%over image columns.\footnote{Only the marginal likelihoods (or unary
%potentials in the language of energy functions) decompose in this
%way. Our prior permits strong coupling across widely separated image
%columns.} We will then present generative models for photometric,
%multiple--view, and 3D features that give rise to MAP formulae that
%can be expressed in precisely this form. This will allow us
%in \sectref{inference} to provide a concise dynamic programming
%solution in the abstract setting of payoff matrices.

General Manhattan environments have structural surfaces oriented in
three cardinal orientations. \textit{Indoor} Manhattan environments
are a special case that consist of a floor plane, a parallel ceiling
plane, and a set of vertical walls extending between them. Each wall
extends all the way from the floor to ceiling, and walls meet at
vertical edges. We always consider environments observed from a camera
located between the floor and ceiling. Since each wall extends from
floor to ceiling, indoor Manhattan environments always project
as a linear chain of walls in the image, as shown in
\figref{ideal-models}. Further, the edges at which adjacent walls meet
can be categorized as concave, convex, or occluding, as illustrated in
\figref{corner-types} and discussed further in \cite{Lee09}.

We assume that vanishing points for the three Manhattan directions are
given. We use the vanishing point detector described by Zhang \etal
\cite{Zhang02} in the monocular setting and that of \cite{FlintCVPR10} in
the multiple view setting. It will greatly simplify the remainder of
this paper if we can assume that vertical lines in the world appear
vertical in the image. To this end we apply the simple rectification
procedure of \cite{FlintCVPR10}.
%\footnote{Note that we could avoid the
%  warping artifacts introduced by this operation and instead carry
%  through geometric calculations assuming a general camera pose, but
%  this would make the remainder of this paper much harder to follow.}

We now describe our parametrization for indoor Manhattan models. Let
the image dimensions be $\Width \times \Height$. Following
rectification, the vertical seams at which adjacent walls meet project
to vertical lines, so each image column intersects exactly one wall
segment. Let the top and bottom of the wall in column $x$ be
$\PfAtX=(x,\YfAtX)$ and $\PcAtX=(x,\YcAtX)$ respectively (depicted in
\figref{corner-types}). Since each $\PfAtX$ lies on the floor plane
and each $\PcAtX$ lies on the ceiling plane, we have
\begin{equation}
\label{eq:manhattan-homology}
  \PfAtX = \Hcf \PcAtX ~.
\end{equation}
where $\Hcf$ is a planar homology \cite{Criminisi01}. We show how to
recover $\Hcf$ in \sectref{floorceil}. Once $\Hcf$ is known, any
indoor Manhattan model is fully described by the values $\{\YfAtX\}$,
leading to the simple parametrization,
\begin{equation}
  \Model = \{ \YfAtX \}_{x=1}^{\Width} ~.
\end{equation}
We query this parametrization as follows. To check whether a pixel
$(x_0,y_0)$ lies on a vertical or horizontal surface we simply need to
check whether $y_0$ is between $\YfAtXX$ and $\YcAtXX$. If we know the
3D position of the floor and ceiling planes then we can recover the
depth of every pixel as follows. If the pixel lies on the floor or
ceiling then we simply back--project a ray onto the corresponding
plane. If not, we back--project onto the vertical plane defining the
wall at that column (the depth of which we can recover from
$\YfAtXX$). Note in particular that the orientation and depth of a
pixel can be recovered from just the floor/wall intersection in its
column; this will be important in later sections.

We now turn to the optimization framework that each subsequent section
will feed into. Let $\{c_i\}$ index the columns at which neighbouring
walls meet in $\Model$. We define the payoff for $\Model$ as
\begin{equation}
  \label{eq:payoffs}
  \ModelPayoff(\Model) = 
  \sum_{x=1}^{\Width} \PixelPayoff(x,\YfAtX) -
  \sum_i \CornerPenalty(c_i) \,
\end{equation}
where the payoff matrix $\PixelPayoff$ assigns payoffs for models with
floor/wall intersections that pass through each pixel, and
$\CornerPenalty$ is a per--corner regulariser which penalizes complex
models. Note that the value of $\PixelPayoff(x,y)$ is \textit{not}
restricted to dependence on pixel $(x,y)$, nor even to a local region
about that pixel; indeed, the payoff functions described in the
following sections incorporate image evidence from widely separated
image regions.

\subsection{Monocular features}

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.35\textwidth]{figures/monocular-gm}
  \caption{The graphical model relating building structures $M$ to
    monocular image features $\Feature$. $\Pixel=(x,y)$ is a
    pixel location and $a$ is the orientation predicted (deterministically) by
    $\Model$ at $\Pixel$.}
  \label{fig:monocular-gm}
\end{figure}

To infer indoor Manhattan models from monocular images we assume the
graphical model shown in \figref{monocular-gm}. We turn first to the
prior $P(\Model~|~\Penalties)$. For a model with $n_1$ concave
corners, $n_2$ convex corners, and $n_3$ occluding corners
(\cf \figref{corner-types}), our prior on models is
\begin{eqnarray}
  \label{eq:model-prior}
  P(\Model ~|~ \Penalties) &=& \frac{1}{Z} 
    {\PenaltyConc}^{n_1} {\PenaltyConv}^{n_2} {\PenaltyOccl}^{n_3}
  % \\
  %Z &=& \sum_{n_c,n_n,n_o} P(\Model ~|~ \Penalties) \\
  %Z&=& \frac{1}{(1-\PenaltyConc)(1-\PenaltyConv)(1-\PenaltyOccl)}~,
\end{eqnarray}
which corresponds to a fixed probability for ``events'' corresponding
to each type of corner and penalizes models for additional
complexity. $Z$ is a normalizing constant.

Our model includes hidden orientation variables
$\Orient_i\in\{1,2,3\}$ for each pixel, with values corresponding to
the three Manhattan orientations (shown as red, green, and blue
regions in \figref{ideal-models}). As described in \sectref{model},
$\Orient$ is deterministic given the model $\Model$. We assume a
linear likelihood for pixel features $\Feature$,
\begin{equation}
\label{eq:feature-likelihood}
  P(\Feature ~|~ \Orient) = \frac{{\PixelModel_\Orient}^T\Feature}
  {\sum_j{\PixelModel_\Orient}^T\Feature_j} ~.
\end{equation}

We now derive MAP inference. The posterior on $\Model$ is
\begin{align}
  \label{eq:mono-post}
  P(\Model ~|~ \Features) = {} &
  \eta P(\Model) \prod_i P(\Feature_i ~|~ \Orient_i^*)
\end{align}
where $\Orient_i^*$ is the orientation deterministically predicted by
model $\Model$ at pixel $\Pixel_i$ and $\eta$ is a normalizing
constant. We have omitted $P(\Orient_i~|~\Model)$ since it equals
1 for $\Orient_i^*$ and 0 otherwise. Taking logarithms,
\begin{align}
  \label{eq:mono-logpost}
  \begin{split}
    \log P(\Model ~|~ \Features) = {} &
    n_1 \PenaltyConc' + n_2 \PenaltyConv' + n_3 \PenaltyOccl'
    \\    
    & + \sum_i \log P(\Feature_i ~|~ \Orient_i^*) + k
  \end{split}
\end{align}
where $\PenaltyOccl'=\log \PenaltyOccl$ and similarly for the other
penalties, and $k$ corresponds to the normalizing denominators in
\eqnref{mono-post} and \eqnref{model-prior}, which we henceforth drop
since it makes no difference to the optimization to come. We can now
put \eqnref{mono-logpost} into payoff form \eqnref{payoffs} by writing
\begin{equation}
  \label{eq:mono-payoffs}
  \begin{split}
    \MonoPayoff(x,\ModelAtX) =&
    \sum_{y'} \log P(\Feature_i ~|~ \Orient_i^*) \\
    \MonoCornerPenalty(c) =& -\lambda_c
  \end{split}
\end{equation}
where $\lambda_c$ is one of $\PenaltyConc$, $\PenaltyConv$, or
$\PenaltyOccl$ according to the category of corner $c$. We show how to
maximize payoffs of this form in \sectref{inference}, which will allow
us to solve MAP inference.

\subsection{Multiple view features}
\begin{figure}[tb]
  \centering \includegraphics[width=0.4\textwidth]{figures/backproject}
  \caption{Pixel correspondences across multiple views are computed by
    back--projection onto the model $\Model$ followed by re--projection into
    auxiliary views.}
  \label{fig:backproject}
\end{figure}
We now formulate the payoff function $\StereoPayoff$ for the case that
multiple views of the scene are available. We assume one base frame
$I_0$ and $M$ auxiliary frames $I_1,\ldots,I_M$. We assume that poses
are given for each camera, as output for example by a
structure--from--motion system, and that cameras are calibrated. We
normalize images intensities to zero mean and unit variance.

Intuitively, we treat inference in this settings as follows. We
consider models $\Model$ in terms of their projection into $I_0$. We
explained in \sectref{model} that models parametrized in image
coordinates specify unique 3D models. Any hypothesized model can
therefore be re--projected into auxiliary frames, giving pixel--wise
correspondences between frames as shown in \figref{backproject}. From
this we compute a photo--consistency measure $\pc(\cdot)$, which
provides the likelihood $P(\StereoData ~|~ \Model)$. The prior remains
as in \eqnref{model-prior}.

Optimizing over photo--consistency has been standard in the stereo
literature for several decades \cite{Scharstein01}; our contribution
is to show that (i) in the particular case of indoor Manhattan models,
photo--consistency can be expressed as a payoff matrix; (ii) that we
can therefore perform efficient and exact global optimization; and
(iii) that this fits naturally within a Bayesian framework alongside
monocular and 3D features.

% SPACE: remove para below
Our approach could also be cast as solving the general stereo problem
where in place of priors based on various pixel--wise norms, our prior
assigns zero probability to all non--indoor--Manhattan
reconstructions.

Let $\reproj_k(\Pixel;\Model)$ be the re--projection of pixel $\Pixel$
from the base frame $I_0$ into auxiliary frame $I_k$ via model
$\Model$. Then
\begin{equation}
  \label{eq:stereo-likelihood}
  \log P(\StereoData ~|~ \Model) = 
   \sum_{\Pixel\in I_o} \sum_{k=1}^M
  \pc(\Pixel,\reproj_k(\Pixel,\Model)) ~,
\end{equation}
where in our experiments $\pc(\Pixel,\Qixel)$ is the sum of squared
differences between pixels $\Pixel$ and $\Qixel$.

%Equation
%\eqnref{stereo-likelihood} could be derived from a graphical model
%assuming a joint Gaussian relationship across correspondences, but we
%have skipped this because it is standard in the literature
%\cite{Scharstein01}.

We explained in \sectref{model} that the depth of each pixel can be
recovered from the location of the floor/wall intersection $\ModelAtX$
in column $x$. Hence we can replace $\reproj_k(\Pixel;\Model)$ with
$\reproj_k(\Pixel;\ModelAtX)$ and write
\begin{equation}
  \label{eq:stereo-payoffs}
  \StereoPayoff(x,y_x) = 
    \sum_{y=1}^{\Height} \sum_{k=1}^M
      \pc(\Pixel,\reproj_k(\Pixel,y_x)) ~,
\end{equation}
where $\Pixel=(x,y)$. To see this, substitute \eqnref{stereo-payoffs}
into \eqnref{payoffs} and observe that the result is
precisely \eqnref{stereo-likelihood}.

Note that the column--wise decomposition \eqnref{stereo-payoffs}
neither commits us to optimizing over columns independently, nor to
ignoring interactions between columns. Such interactions come into
effect when we optimize over the full payoff matrix
in \sectref{inference}, and our results will show that widely
separated image regions often interact strongly. The derivations
in this section follow deductively from the indoor Manhattan
assumption; the only approximation is the following.

\textbf{Occlusions}. We have ignored self--occlusions in
\eqnref{stereo-likelihood}. For short baselines (such as frames
sampled over a few seconds from a moving camera), this is
unproblematic since indoor environments tend to be mostly convex from
any single point of view. Even in highly non--convex environments our
system achieves excellent results by integrating 3D and monocular
features, and enforcing strong global consistency, as will be shown in
\sectref{results}.

%\subsubsection{Extension: using orientation information}

%\subsubsection{EM for Occlusion Resolution}

\subsection{3D features}
\begin{figure}[tb]
  \centering
  \includegraphics[width=0.35\textwidth]{figures/3d-gm}
  \caption{The graphical model relating indoor Manhattan models to 3D
    points. The hidden variable $t$ indicates whether the point is
    inside, outside, or coincident with the model.}
  \label{fig:3d-gm}
\end{figure}

In this section we explore the context in which a 3D point cloud is
available during inference. The point clouds generated by
structure--from--motion systems are typically too sparse for direct
reconstruction, but can provide useful cues alongside monocular and
stereo data.

\begin{figure}[tb]
  \centering \includegraphics[width=0.3\textwidth]{figures/point-projs}
  \caption{Depth measurements $d_i$ might be generated by a surface in
    our model (represented by $t_i=\ON$) or by an object inside or
    outside the environment (in which case $t_i=\IN,\OUT$
    respectively).}
  \label{fig:point-projs}
\end{figure}

%For each pixel $\Pixel_i$ a depth measurement $d_i$ is made. The
%relationship between $d$ and the model $M$ is determined by a hidden
%variable $t\in\{\textsf{IN},\textsf{ON},\textsf{OUT}\}$, which is
%interpreted as follows. Let $\DepthAtPGivenModel$ be the true depth of
%the model $\Model$ in the direction $\Pixel$. If $t=\textsf{IN}$ then
%$\Depth$ corresponds to an object inside the room,

Our graphical model for 3D data is depicted in \figref{3d-gm}. The
model $\Model$ is sampled according to the prior \eqnref{model-prior},
then depth measurements $d_i$ are generated for pixels
$\Pixel_i$. Many such measurements will correspond to clutter or
measurement errors, rather than to the walls represented by
$\Model$. Our model captures this uncertainty explicitly through
the latent variable $t_i$, which has following interpretation. If
$t_i=\ON$ then $d_i$ corresponds to some surface represented
explicitly in $\Model$. Otherwise, either $t_i=\IN$, meaning some
clutter object within the room was measured, or $t_i=\OUT$, in which
case an object outside the room was measured, such as through a
window. The likelihoods we use are
\begin{align}
  \label{eq:x-inside}
  P(\Depth~|~\Pixel,\Model,\textsf{IN}) &=
  \begin{cases}
    \alpha, & \mbox{if } 0 < \Depth < \DepthAtPGivenModel \\
    0, & \mbox{otherwise} \\
  \end{cases} \\
  \label{eq:x-outside}
  P(d~|~\Pixel,\Model,\textsf{OUT}) &=
  \begin{cases}
    \beta~, & \mbox{if } \DepthAtPGivenModel < \Depth < \Dmax \\
    0~, & \mbox{otherwise} \\
  \end{cases} \\
  P(d ~|~ \Pixel,\Model,\textsf{ON}) &=
  \mathcal{N}(d~;~\DepthAtPGivenModel,\sigma) ~.
\end{align}
%If $t=\textsf{OUT}$ then $\Depth$ corresponds to a point outside
%the boundary of the room, perhaps observed through a window or
%doorway,
%\begin{align}
%  \label{eq:x-outside}
%  P(d~|~\Pixel,\Model,\textsf{OUT}) &=
%  \begin{cases}
%    \beta~, & \mbox{if } \DepthAtPGivenModel < \Depth < \Dmax \\
%    0~, & \mbox{otherwise} \\
%  \end{cases} ~.
%\end{align}
%Finally, if $t=\textsf{ON}$ then $\Depth$ corresponds to a surface
%on $M$,
%\begin{align}
%  P(d ~|~ \Pixel,\Model,\textsf{ON}) &=
%  \mathcal{N}(d~;~\DepthAtPGivenModel,\sigma)  ~.
%\end{align}
where $\alpha$ and $\beta$ are determined by the requirement that the
probabilities sum to $1$ and $\DepthAtPGivenModel$ denotes the depth
predicted by $\Model$ at $\Pixel$. We compute likelihoods on $\Depth$ by
marginalizing,
\begin{align}
  \label{eq:3d-likelihood}
  P(\Depth~|~\Pixel,\Model) &=
   \sum_{t} P(\Depth~|~\Pixel,\Model,t) P(t)
  ~,
\end{align}
%Now \eqnref{3d-full-posterior} is a sum over three possible values for
%$t_i$ and we can simplify this expression by distinguishing the case
%where $d_i>\DepthAtPGivenModel$, in which case \eqnref{x-inside} is
%$0$, from the case where the opposite is true, in which case
%\eqnref{x-outside} is $0$. Hence,
%\begin{align}
%  \label{eq:3d-likelihood}
%  P(\Depth_i~|~\Pixel_i,\Model) &= \alpha +
%  \tON \mathcal{N}(d_o~;~\DepthAtPGivenModel,\sigma)
%  \\
%  \alpha &=
%  \begin{cases}
%    \frac{\tIN}{\DepthAtPGivenModel}, & \mbox{if } d_i < \DepthAtPGivenModel \\
%    \frac{\tOUT}{\Dmax-\DepthAtPGivenModel}, & \mbox{otherwise} \\
%  \end{cases}~.
%\end{align}
where the prior $P(t)$ is a look--up table with three entries denoted
$\tIN$, $\tOUT$, and $\tON$.

As explained in \sectref{model}, computing the depth of the model at
pixel $\Pixel$ requires knowledge only of the floor/wall intersection
$\ModelAtX$ in column $x$, so we substitute
\begin{equation}
  P(\Depth ~|~ \Pixel, \ModelAtX) =
  P(\Depth ~|~ \Pixel, \Model) ~.
\end{equation}
Let $\Depths$ denote all depth measurements, $\Pixels$ denote all pixels, and $\mathcal{D}_x$ contain
indices for all depth measurements in column $x$. Then
\begin{align}
  P(\Model ~|~ \Depths,\Pixels) = & 
  P(\Model) \prod_{x} \prod_{i\in\mathcal{D}_x} P(\Depth_i~|~\Pixel_i,\ModelAtX)
  \\
  \log P(\Model ~|~ \Depths,\Pixels) = & 
  P(M) + \sum_{x} \Bigl( \sum_{i\in\mathcal{D}_x}
  \log P(\Depth_i~|~\Pixel_i,\ModelAtX) \Bigr) ~,
\end{align}
which we write in payoff form as
\begin{equation}
  \DepthPayoff(x,\ModelAtX) = 
  \sum_{i\in\mathcal{D}_x} \log P(\Depth_i~|~\Pixel_i,\ModelAtX)
\end{equation}
and the penalty function $\CornerPenalty$ remains as in \eqnref{mono-payoffs}.

\subsection{Combining features}
We combine photometric, stereo, and 3D data into a joint model by
assuming conditional independence given $M$,
\begin{equation}
  \begin{split}
    P&(M ~|~ X_{\textsf{mono}}, X_{\textsf{stereo}}, X_{\textsf{3D}}) = \\
    & P(M) 
      P(X_{\textsf{mono}} ~|~ M)
      P(X_{\textsf{stereo}} ~|~ M)
      P(X_{\textsf{3D}} ~|~ M)
  \end{split}
\end{equation}
Taking logarithms leads to summation over payoffs,
\begin{equation}
  \JointPayoff(\vect{x}) =
  \MonoPayoff(\vect{x}) + 
  \StereoPayoff(\vect{x}) +
  \DepthPayoff(\vect{x}) ~.
\end{equation}

\subsection{Resolving the floor and ceiling planes}
\label{sect:floorceil}
We resolve the equation of the floor and ceiling planes as follows. If
$C$ is the camera matrix for any frame and $\vvpt$ is the vertical
vanishing in that frame, then $\vect{n} = C^{-1}\vvpt$ is normal to
the floor and ceiling planes. We sweep a plane with this orientation
through the scene, recording at each step the number of points within
a distance $\delta$ of the plane ($\delta$=0.1\% of the diameter of
the point cloud in our experiments). We take as the floor and ceiling
planes the minimum and maximum locations such that the plane contains
at least $5$ points. We found that this simple heuristic worked
without failure on our training set.

Let the two non--vertical vanishing points be $\lvpt$ and $\rvpt$ and
let $\vect{h}=\lvpt\times\rvpt$. Select any two corresponding points
$\vect{x_f}$ and $\vect{x_c}$ on the floor and ceiling planes
respectively. Then the Manhattan homology defined in \eqnref{manhattan-homology}
is given by
\begin{equation}
  \Hcf = I + \mu\frac{\vvpt\vect{h}^T}{\vvpt \cdot \vect{h}} ~,
\end{equation}
where $\mu = <\vvpt,\vect{x_c},\vect{x_f},
\vect{x_c}\times\vect{x_f}\times\vect{h}>$ is the characteristic cross
ratio of $\Hcf$.

\section{Inference}
\label{sect:inference}
We have reduced MAP inference to optimization over a payoff matrix:
\begin{equation}
  \EstModel = \argmax_\Model\limits \sum_x \PixelPayoff(x,\ModelAtX) -
  \sum_i \CornerPenalty(c_i) \,
\end{equation}
In previous work \cite{FlintECCV10} we showed that if an indoor
Manhattan model $\PartialModel$ is optimal over image columns
$\left[1,\PartialCol\right]$, then the ``cropped'' model
$\CroppedModel$, obtained by restricting $\PartialModel$ to the
sub--interval $\left[1,\CroppedCol\right] ~\CroppedCol<\PartialCol$,
must itself be optimal over that sub--interval. This permits a dynamic
programming solution in which $\EstModel$ is built up from left to
right.

Our algorithm differs from that of \cite{FlintECCV10} in the following
respects. First, we optimize over general payoff matrices of the form
\eqnref{payoffs}; whereas neither $\StereoPayoff$ nor $\DepthPayoff$
decomposes as assumed in \cite{FlintECCV10}.
Second, we do not include the number of
corners as a state variable, but instead accumulate penalties directly
into the objective function, which reduces complexity by $O(K)$ where
$K$ is the number of walls in the model. For completeness we give
revised recurrence relations in an appendix.

\section{Training}
In this section we address the learning of model parameters from
labeled training data. We turn first to the parameter $\PixelModel$
relating photometric features to the orientation variables $\Orient_i$
via \eqnref{feature-likelihood}. We employ the following bootstrapping
algorithm to learn $\PixelModel$. We begin by sampling $k$ pixels at
random from the images in the training set and use these to train a
classifier (in our case a multi--class SVM with three classes) for the
task of mapping pixel features $\Feature$ to orientations
$\Orient$. We then run the complete inference procedure on the entire
training set, using the current classifier to evaluate the
log--likelihood \eqnref{mono-payoffs}. We compute a pixel--wise loss
$\PixelLoss(\EstModel,\TrueModel)$ with respect to ground truth. In
our experiments $\PixelLoss$ is the relative depth error,
\begin{equation}
  \label{eq:loss}
  \PixelLoss(\EstModel,\TrueModel) = 
  \Bigl| \frac{\DepthAt(\vect{p};\EstModel) - \DepthAt(\vect{p};\TrueModel)}
    {\DepthAt(\vect{p};\TrueModel)}\Bigr| ~.
\end{equation}
We then sample $k$ additional pixels to be added to the training set,
where each pixel is selected with probability proportional to the loss
\eqnref{loss}, then re--train the pixel classifier and repeat to
convergence. That is, we add pixels at which \textit{the image--level
  inference procedure} is making the greatest mistakes, which biases
learning towards portions of the training set where the inference
process as a whole, rather than the pixel--level classifier, is making
mistakes.

\textbf{Model prior parameters $\Penalties$}. We assign a beta distribution
with $\alpha=\beta=1$ as hyper--prior for $\Penalties$. MAP estimates
are then given by
\begin{equation}
  \hat{\lambda_k} = \frac{\expect[n_k]}{\expect[n_k]+1}
  % this makes sense as 0 <= lambda_k < 1 so 0 <= -log(lambda_k) <= inf
\end{equation}
where expectations are over the training set.

\textbf{3D indicators}. We assign a uniform hyper--prior to all
$\tparams$ representing a valid probability distribution (\ie positive
and unit sum), then perform gradient descent on the posterior
$P(\tparams~|~\{\Model,\{p_i,x_i\}\})$.

\newcommand{\Sample}[2]{\includegraphics[width=0.16\textwidth]{figures/learning/iter#1_frame025_#2.png}}

\begin{figure}[tb]%
  \centering
  \begin{tabular}{ccc}
    \vspace{2mm} & \textbf{Training samples} & \textbf{MAP model} \\
    \rotatebox{90}{\hspace{4mm}Iteration 1} & \Sample{000}{active} & \Sample{000}{dp} \\
    \rotatebox{90}{\hspace{4mm}Iteration 3} & \Sample{002}{active} & \Sample{002}{dp} \\
    \rotatebox{90}{\hspace{4mm}Iteration 5} & \Sample{005}{active} & \Sample{005}{dp} \\
    \rotatebox{90}{\hspace{3mm}Iteration 10} & \Sample{016}{active} & \Sample{016}{dp} \\
    \rotatebox{90}{\hspace{3mm}Iteration 30} & \Sample{029}{active} & \Sample{029}{dp} \\
  \end{tabular}
  \caption{Snapshots of our bootstrap learning algorithm. The left
    column the pixels that the SVM was trained on in each iteration
    and the right column shows the corresponding MAP model. Each
    iteration injects incorrect pixels into the training set, leading
    to a concentration about surface boundaries since these locations
    are the most often confused by our model. This corresponds to
    the intuition that pixels near surface boundaries are the most
    ``important'' for the SVM to correctly classify since our model
    will leverage global consistency to ``fill in the blanks'' in
    other regions.}
  \label{fig:learning}
\end{figure}

\section{Results}\label{sect:results}
Our data--set consists of 18 manually annotated video sequences of
indoor scenes averaging 59 seconds in duration. We sample frames at
one second intervals and divide frames into consecutive groups of 3
(one base frame and two auxiliary frames). Our training set consists of
150 such triplets generated from 8 different sequences. Our test set
contains 204 triplets from the remaining 10 sequences. No sequence
appears in both the training and test sets.

To acquire ground truth data we reconstructed camera trajectories
using structure--from--motion software (we use the PTAM system of
Klein and Murray \cite{Klein07}) then manually specify the ground
truth floor--plan. Recall that we seek to recover the
\textit{boundaries} of the environment, whether or not they are
visible at every point. When our algorithm ignores clutter within a
room, we consider that a \textit{success}.

The monocular features $\Feature_i$ consist of 3 RGB channels, 3 HSV
channels, 24 Gabor filters (4 scales, 6 orientations), and 3 binary
line sweep features \cite{Lee09}. For stereo we use patches of size
$5 \times 5$.

We compute two error metrics: the labeling accuracy, which is the
proportion of all pixels that were labeled with the correct
orientation, and the mean relative depth error \eqnref{loss}. While
the latter better captures similarity to the ground truth, not all the
systems we compare against have direct 3D interpretations and in such
cases we must compare on labeling accuracy.

To the best of our knowledge, there is no previously published work on
precisely this problem (indoor--Manhattan reasoning from multiple
views) so we compare with two alternative systems, though neither
comparison is ideal.

Our first comparison is with the approach of Brostow \etal
\cite{Brostow08}, who performed semantic segmentation by training a
per--pixel classifier on structure--from--motion cues. Our
implementation of their system uses exactly the features they
describe, with classes corresponding to the three Manhattan
orientations. While they trained a randomized forest, we trained a
multi--class SVM because a reliable SVM library was more readily
available to us. Given the margin between our results it is
unlikely that a different classifier would significantly change the
outcome.

The second comparison is with the monocular approach of
Lee \etal \cite{Lee09}. One would of course expect a multiple view
approach to outperform a monocular approach, but as one of the very
few previous approaches to have explicitly leveraged the indoor
Manhattan assumption we feel this comparison is important to
demonstrate the benefit of a Bayesian framework and integration of
stereo and 3D cues.

The performance of each system is shown in \figref{performance}. Our
system significantly out--performs both others. Even when restricted
to monocular features, our system outperforms \cite{Brostow08}, which
has access to 3D cues. This reflects the utility of global consistency
and the indoor Manhattan representation in our approach.

The initialization procedure of \cite{Lee09} fails for 31\% of our
training images, so at the bottom of \figref{performance} we show
results for their system after excluding these images. Labeling
accuracy increases to within 3\% of our monocular--only results,
though on the depth error metric a margin of 10\% remains. This
illustrates the effect of our training procedure, which optimizes for the
depth error.

\Figref{performance} also shows that joint estimation
is superior to using any one sensor modality alone. Anecdotally we
find that using 3D cues alone often fails within large textureless
regions in which the structure--from--motion system failed to track
any points, whereas stereo or monocular cues alone often perform
better in such regions but can lack precision at corners and
boundaries.

\Figref{timing} shows timing results for our system. For each triplet
of frames, our system requires on average less than one second to
compute features for all three frames and less than 100 milliseconds to perform
optimization. 
%Computation of stereo photoconsistency dominates
%efficieny.

\begin{figure}[tb]
  \centering
  \begin{tabular}{@{}lp{2.1cm}p{1.8cm}@{}}
    \toprule
    Algorithm & Mean depth error (\%) & Labeling accuracy (\%) \\
    \midrule
    Our approach (full) & \textbf{14.5} & \textbf{75.5} \\
    \hspace{1mm} Stereo only & 17.4 & 69.5 \\
    \hspace{1mm} 3D only & 15.2 & 71.1 \\
    \hspace{1mm} Monocular only & 24.8 & 69.2 \\
    Brostow \etal \cite{Brostow08} && 40.6  \\  % Tue_Brostow_NoNormals_mcSVM
    Lee \etal \cite{Lee09} & 79.8 & 45.5 \\
    \hspace{1mm}excluding failures\footnotemark & 34.1 & 66.2 \\
    \bottomrule
  \end{tabular}
  \vspace{0.2cm}
  \caption{Performance on our data--set. Labeling accuracy is the
    percentage of correctly labeled pixels over the data--set, and
    depth error is a per--pixel average of \eqnref{loss}.}
  \label{fig:performance}
\end{figure}
\footnotetext{This row excludes cases for which \cite{Lee09}
  was unable to find overlapping lines during initialization.}

\newcommand{\Res}[4]{\includegraphics[width=0.15\textwidth]{figures/full_results/#1/#2_#3_frame#4_dp.png}}

\newcommand{\TopRes}[3]{\Res{top}{#1}{#2}{#3}}
\newcommand{\MedRes}[3]{\Res{median}{#1}{#2}{#3}}
\newcommand{\FailRes}[3]{\Res{fail}{#1}{#2}{#3}}

\begin{figure*}[tb]%
  \centering
  \begin{tabular}{ccc}
    \textbf{Results above 90th percentile} &
    \textbf{Results near median} &
    \textbf{Failures (below 10th percentile)} \\

    \TopRes{lab}{foyer2}{046}
    \TopRes{lab}{foyer1}{005} &

    \MedRes{exeter}{bursary}{008}
    \MedRes{exeter}{mcr1}{015} &

    \FailRes{exeter}{bursary}{021}
    \FailRes{exeter}{mcr1}{029} \\

    \TopRes{exeter}{mcr1}{024}
    \TopRes{lab}{foyer2}{001} &

    \MedRes{exeter}{mcr1}{021}
    \MedRes{exeter}{mcr1}{042} &

    \FailRes{exeter}{mcr1}{039}
    \FailRes{lab}{kitchen1}{017} \\

    \TopRes{lab}{foyer2}{035}
    \TopRes{som}{corr1}{013} &

    \MedRes{lab}{kitchen1}{091}
    \MedRes{exeter}{mcr1}{049} &

    \FailRes{lab}{kitchen1}{089}
    \FailRes{som}{corr1}{006} \\

    %\TopRes{lab}{foyer2}{046} &
    %\MedRes{lab}{ground1}{010} &
    %\FailRes{som}{corr1}{020} \\
  \end{tabular}
  \caption{Models output from our system. The left column shows
  results above the 90th percentile of performance (relative depth
  error), the middle column shows results near median performance, and
  the right column shows failure cases.}
  \label{fig:results-pics}
\end{figure*}

\section{Conclusion}
We have presented a Bayesian framework for scene understanding in the
context of a moving camera. Our approach draws on the indoor Manhattan
assumption introduced for monocular reasoning and we have shown that
techniques from monocular and stereo vision can be integrated with 3D
data in a coherent Bayesian framework.

In future work we intend to use indoor Manhattan models to reason
about objects, actions, and scene categories. We also intend to
investigate structural SVMs for learning parameters, which may allow
us to relax the conditional independence assumptions between sensor
modalities.

\section{Appendix}\label{sect:appendix}
\textbf{Recurrence relations for MAP inference.} Let
$\fOUT(x,y,\Orient)$, $1 \leq x \leq \Width$, $1 \leq y \leq \Height$,
$\Orient \in \{1,2\}$ be the maximum payoff for any indoor Manhattan
model $\Model$ spanning columns $\left[1, x\right]$, such that (i) $M$
contains a floor/wall intersection at $(x,y)$, and (ii) the wall that
intersects column $x$ has orientation $a$. Then $\fOUT$ can be
computed by recursive evaluation of the recurrence relations,
\begin{align}
  \label{eq:fout-recurrence}
    \fOUT(x,y,a) = {} & \max_{a'\in\{1,2\}}
      \begin{cases}
        \fUP(x,y-1,a') - \gamma(x) & \\
        \fDOWN(x,y+1,a') - \gamma(x) & \\
        \fIN(x,y,a') - \gamma(x) &
      \end{cases}\\
  \label{eq:fup-recurrence}
  \fUP(x,y,a) = {} & \max \Bigl(\fIN(\cdot), \fUP(x,y-1,a)\Bigr)~,\\
  \label{eq:fdown-recurrence}
  \fDOWN(x,y,a) = {} & \max \Bigl(\fIN(\cdot), \fDOWN(x,y+1,a)\Bigr)~,\\
  \label{eq:fin-recurrence-basic}
  \fIN(x,y,a) = {} &
  \max_{x'<x}\Bigl(\fOUT(x',y',a)+ \Delta\Bigr) ~,\\
  & \Delta = \sum_{i=x'}^{x} \PixelPayoff(i,y') ~.
\end{align}
Here we have treated $\fIN$, $\fUP$, and $\fDOWN$ simply as notational
placeholders; for their interpretations in terms of sub--problems see
\cite{FlintECCV10}. Finally, the base cases are
\begin{align}
  \fOUT(0,y,a) = {} & 0 \hspace{1cm} \forall y,a \\
  \fUP(x,0,a) = {} & \infty \hspace{1cm} \forall x,a \\
  \fDOWN(x,\Width,a) = {} & \infty \hspace{1cm} \forall x,a ~.
\end{align}

\begin{figure}[tb]
  \centering
  \begin{tabular}{@{}lp{2.1cm}p{1.8cm}@{}}
    \toprule
    Component & Time (ms) & stddev (ms) \\
    \midrule
    Monocular features & 160 & 7.6 \\
    Stereo features & 730 & 43 \\
    3D features & 8.8 & 0.05 \\
    Optimization & 102 & 15 \\
    \textbf{Total} & \textbf{997} & \textbf{43} \\
    \bottomrule
  \end{tabular}
  \vspace{0.2cm}
  \caption{Timing results for our system, averaged over the test
    set. Times show complete processing time for each triplet of
    frames (base frame plus two auxiliary frames).}
  \label{fig:timing}
\end{figure}

{\small
\bibliographystyle{ieee}
\bibliography{AVLstrings,VisionRefs}
}

### 2014-07-24 iccv2011/rebuttal.txt
move papers/* to top level

We thank all reviewers for encouraging and constructive reviews. We are pleased that all reviewers offered broad encouragement of our approach and noted general satisfaction with our results. We are confident that we can address all suggestions and clarification requests in the camera-ready manuscript.

R1 requested several clarifications, each of which we have addressed in the final paper; we summarise the answers below:
1) Regarding occlusions- we assume locally convex environments (lines 408-417); this is violated surprisingly rarely since we are always interested in the boundaries of the environment and never the objects or clutter within. This means the DP optimization simply ignores occlusions. When we do encounter non-convex environments we find that global optimization together with a prior on model complexity (eq. 4) produces pleasing results regardless.
2) Features- we perform intensity and contrast normalization but we do not handle specularities or other non-linear lighting phenomena. Our intention is to show that models from stereo vision fit naturally into our framework; fully leveraging the best modern stereo algorithms remains future work.
3) Training- We trained and evaluated on separate sets of sequences. No place or room observed in the training set was also present in the test set.
4) Inference- All our experiments were performed with 2 auxiliary frames, which were sampled 1 second before and after each base frame. In the video sequence we selected base frames at 8-second intervals, and at each step we discarded the previous model and re-estimated. Each video frame shows the most recent model re-projected into the current frame.
5) Timing- We include further timing results in the final paper in line with the request by R2. Training on 150 frames takes roughly 15 minutes.

We thank R2 for two suggested citations. The work in (Lee et al, 2010) is relevant to ours primarily via earlier work (Lee et al, CVPR 2009), which our paper cited. We have added a citation for (Hedau et al, 2009), though our paper overlaps only marginally with this monocular approach. We regret that R2 found Section 3 unclear and we will revise extensively for the final submission. We have added timing results to the final paper as requested; we regret that this was omitted from the review manuscript.

We agree with R3 regarding overlap with [7]. We intended this paper to be understandable on its own since [7] was published less than a year ago and is not widely known, so we compromised by repeating some of their results during lines 530-553. We were aware of Erick Delage's early work in this area and regret omitting that citation; we have rectified this in our final submission. Regarding comparison with [7], the "Monocular only" line in Figure 8 is identical to the system of [7]; our final paper states this explicitly. (Compared to [7], the addition of stereo and 3D features reduces mean depth error by 10%.)

### 2014-07-24 eccv2010/eccv2010submission.tex
move papers/* to top level

% last updated in April 2002 by Antje Endemann
% Based on CVPR 07 and LNCS, with modifications by DAF, AZ and elle, 2008 and AA, 2010

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{lineno}
\usepackage{color}
\begin{document}
\renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
\renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}} 
\linenumbers
\pagestyle{headings}
\mainmatter
\def\ECCV10SubNumber{***}  % Insert your submission number here

\title{Author Guidelines for ECCV Submission} % Replace with your title

\titlerunning{ECCV-10 submission ID \ECCV10SubNumber}

\authorrunning{ECCV-10 submission ID \ECCV10SubNumber}

\author{Anonymous ECCV submission}
\institute{Paper ID \ECCV10SubNumber}

\maketitle

\begin{abstract}
The abstract should summarize the contents of the paper and should
contain at least 70 and at most 300 words. It should be set in 9-point
font size and should be inset 1.0~cm from the right and left margins.
\dots
\end{abstract}

\section{Introduction}

Please follow the steps outlined below when submitting your manuscript.

%-------------------------------------------------------------------------

\subsection{Language}

All manuscripts must be in English.

\subsection{Paper length}
The basic length is 12 pages, but up to two additional pages may be
purchased in the final printed proceedings. This brings the {\em
maximum} length for submission to 14 pages. Overlength papers will
simply not be reviewed. This includes papers where the margins and
formatting are deemed to have been significantly altered from those
laid down by this style guide.  The reason such papers will not be
reviewed is that there is no provision for supervised revisions of
manuscripts. The reviewing process cannot determine the suitability of
the paper for presentation in 14 pages if it is reviewed in 16.

\subsection{Dual submission}

By submitting a manuscript to ECCV, the author(s) assert that it has
not been previously published in substantially similar
form. Furthermore, no paper which contains significant overlap with
the contributions of this paper either has been or will be submitted
during the ECCV 2010 review period to either a journal or a
conference.

If there are any papers that may appear to the reviewers to violate
this condition, then it is your responsibility to (1) cite these
papers (preserving anonymity as described in section~\ref{sec:blind}
of this example paper, (2) argue in the body of your paper why your
ECCV paper is nontrivially different from these concurrent
submissions, and (3) include anonymized versions of those papers in
the supplemental material.

\subsection{Supplemental Material}

Authors may optionally upload supplemental material. Typically, this
material might include videos of results that cannot be included in
the main paper, anonymized related submissions to other conferences
and journals, and appendices or technical reports containing extended
proofs and mathematical derivations that are not essential for
understanding of the paper. Note that the contents of the supplemental
material should be referred to appropriately in the paper and that
reviewers are not obliged to look at it.

All supplemental material must be zipped or tarred into a single
file. There is a 50MB limit on the size of this file. The deadline for
supplemental material is five days after the main paper deadline.

%------------------------------------------------------------------------- 
\subsection{Line numbering}

All lines should be numbered, as in this example document.  This makes
reviewing more efficient, because reviewers can refer to a line on a
page.  If you are preparing a document using a non-\LaTeX\
document preparation system, please arrange for an equivalent line numbering.

\subsection{Mathematics}

Please number all of your sections and displayed equations.  Again,
this makes reviewing more efficient, because reviewers can refer to a
line on a page.  Also, it is important for readers to be able to refer
to any particular equation.  Just because you didn't refer to it in
the text doesn't mean some future reader might not need to refer to
it.  It is cumbersome to have to use circumlocutions like ``the
equation second from the top of page 3 column 1''.  (Note that the
line numbering will not be present in the final copy, so is not an
alternative to equation numbers).  Some authors might benefit from
reading Mermin's description of how to write mathematics:
\url{http://www.cvpr.org/doc/mermin.pdf}.

\section{Blind review}
\label{sec:blind}

Many authors misunderstand the concept of anonymizing for blind
review.  Blind review does not mean that one must remove
citations to one's own work---in fact it is often impossible to
review a paper unless the previous citations are known and
available.

Blind review means that you do not use the words ``my'' or ``our''
when citing previous work.  That is all.  (But see below for 
techreports).

Saying ``this builds on the work of Lucy Smith [1]'' does not say
that you are Lucy Smith, it says that you are building on her
work.  If you are Smith and Jones, do not say ``as we show in
[7]'', say ``as Smith and Jones show in [7]'' and at the end of the
paper, include reference 7 as you would any other cited work.

An example of a bad paper:
\begin{quote}
\begin{center}
    An analysis of the frobnicatable foo filter. 
\end{center}

   In this paper we present a performance analysis of our
   previous paper [1], and show it to be inferior to all
   previously known methods.  Why the previous paper was
   accepted without this analysis is beyond me.

   [1] Removed for blind review
\end{quote}

An example of an excellent paper:

\begin{quote}
\begin{center}
     An analysis of the frobnicatable foo filter. 
\end{center}

   In this paper we present a performance analysis of the
   paper of Smith [1], and show it to be inferior to
   all previously known methods.  Why the previous paper
   was accepted without this analysis is beyond me.

   [1] Smith, L and Jones, C. ``The frobnicatable foo
   filter, a fundamental contribution to human knowledge''.
   Nature 381(12), 1-213.
\end{quote}

If you are making a submission to another conference at the same time,
which covers similar or overlapping material, you may need to refer to that
submission in order to explain the differences, just as you would if you
had previously published related work.  In such cases, include the
anonymized parallel submission~\cite{Authors10} as additional material and
cite it as
\begin{quote}
1. Authors. ``The frobnicatable foo filter'', BMVC 2010 Submission ID 324,
Supplied as additional material {\tt bmvc10.pdf}.
\end{quote}

Finally, you may feel you need to tell the reader that more details can be
found elsewhere, and refer them to a technical report.  For conference
submissions, the paper must stand on its own, and not {\em require} the
reviewer to go to a techreport for further details.  Thus, you may say in
the body of the paper ``further details may be found
in~\cite{Authors10b}''.  Then submit the techreport as additional material.
Again, you may not assume the reviewers will read this material.

Sometimes your paper is about a problem which you tested using a tool which
is widely known to be restricted to a single institution.  For example,
let's say it's 1969, you have solved a key problem on the Apollo lander,
and you believe that the ECCV audience would like to hear about your
solution.  The work is a development of your celebrated 1968 paper entitled
``Zero-g frobnication: How being the only people in the world with access to
the Apollo lander source code makes us a wow at parties'', by Zeus.

You can handle this paper like any other.  Don't write ``We show how to
improve our previous work [Anonymous, 1968].  This time we tested the
algorithm on a lunar lander [name of lander removed for blind review]''.
That would be silly, and would immediately identify the authors. Instead
write the following:
\begin{quotation}
\noindent
   We describe a system for zero-g frobnication.  This
   system is new because it handles the following cases:
   A, B.  Previous systems [Zeus et al. 1968] didn't
   handle case B properly.  Ours handles it by including
   a foo term in the bar integral.

   ...

   The proposed system was integrated with the Apollo
   lunar lander, and went all the way to the moon, don't
   you know.  It displayed the following behaviours
   which show how well we solved cases A and B: ...
\end{quotation}
As you can see, the above text follows standard scientific convention,
reads better than the first version, and does not explicitly name you as
the authors.  A reviewer might think it likely that the new paper was
written by Zeus, but cannot make any decision based on that guess.
He or she would have to be sure that no other authors could have been
contracted to solve problem B. \\

\noindent FAQ: Are acknowledgements OK?  No. Please {\bf omit
acknowledgements} in your review copy; they can go in the final copy.

\section{Manuscript Preparation}
This is an edited version of Springer LNCS instructions adapted for
ECCV 2010 first paper submission.

You are strongly encouraged to use \LaTeX2$_\varepsilon$ for the
preparation of your
camera-ready manuscript together with the corresponding Springer
class file \verb+llncs.cls+.

We would like to stress that the class/style files and the template
should not be manipulated and that the guidelines regarding font sizes
and format should be adhered to. This is to ensure that the end product
is as homogeneous as possible.

\subsection{Printing Area}
The printing area is $122  \; \mbox{mm} \times 193 \;
\mbox{mm}$.
The text should be justified to occupy the full line width,
so that the right margin is not ragged, with words hyphenated as
appropriate. Please fill pages so that the length of the text
is no less than 180~mm.

\subsection{Layout, Typeface, Font Sizes, and Numbering}
Use 10-point type for the name(s) of the author(s) and 9-point type for
the address(es) and the abstract. For the main text, please use 10-point
type and single-line spacing.
We recommend using Computer Modern Roman (CM) fonts, Times, or one
of the similar typefaces widely used in photo-typesetting.
(In these typefaces the letters have serifs, i.e., short endstrokes at
the head and the foot of letters.)
Italic type may be used to emphasize words in running text. Bold
type and underlining should be avoided.
With these sizes, the interline distance should be set so that some 45
lines occur on a full-text page.

\subsubsection{Headings.}

Headings should be capitalized
(i.e., nouns, verbs, and all other words
except articles, prepositions, and conjunctions should be set with an
initial capital) and should,
with the exception of the title, be aligned to the left.
Words joined by a hyphen are subject to a special rule. If the first
word can stand alone, the second word should be capitalized.
The font sizes
are given in Table~\ref{table:headings}.
\setlength{\tabcolsep}{4pt}
\begin{table}
\begin{center}
\caption{Font sizes of headings. Table captions should always be
positioned {\it above} the tables. The final sentence of a table
caption should end without a full stop}
\label{table:headings}
\begin{tabular}{lll}
\hline\noalign{\smallskip}
Heading level & Example & Font size and style\\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
Title (centered)  & {\Large \bf Lecture Notes \dots} & 14 point, bold\\
1st-level heading & {\large \bf 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bf 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bf Headings.} Text follows \dots & 10 point, bold
\\
4th-level heading & {\it Remark.} Text follows \dots & 10 point,
italic\\
\hline
\end{tabular}
\end{center}
\end{table}
\setlength{\tabcolsep}{1.4pt}

Here are
some examples of headings: ``Criteria to Disprove Context-Freeness of
Collage Languages'', ``On Correcting the Intrusion of Tracing
Non-deterministic Programs by Software'', ``A User-Friendly and
Extendable Data Distribution System'', ``Multi-flip Networks:
Parallelizing GenSAT'', ``Self-determinations of Man''.

\subsubsection{Lemmas, Propositions, and Theorems.}

The numbers accorded to lemmas, propositions, and theorems etc. should
appear in consecutive order, starting with the number 1, and not, for
example, with the number 11.

\subsection{Figures and Photographs}
\label{sect:figures}

Please produce your figures electronically and integrate
them into your text file. For \LaTeX\ users we recommend using package
\verb+graphicx+ or the style files \verb+psfig+ or \verb+epsf+.

Check that in line drawings, lines are not
interrupted and have constant width. Grids and details within the
figures must be clearly readable and may not be written one on top of
the other. Line drawings should have a resolution of at least 800 dpi
(preferably 1200 dpi).
For digital halftones 300 dpi is usually sufficient.
The lettering in figures should have a height of 2~mm (10-point type).
Figures should be scaled up or down accordingly.
Please do not use any absolute coordinates in figures.

Figures should be numbered and should have a caption which should
always be positioned {\it under} the figures, in contrast to the caption
belonging to a table, which should always appear {\it above} the table.
Please center the captions between the margins and set them in
9-point type
(Fig.~\ref{fig:example} shows an example).
The distance between text and figure should be about 8~mm, the
distance between figure and caption about 5~mm.
\begin{figure}
\centering
\includegraphics[height=7.2cm]{eijkel2}
\caption{One kernel at $x_s$ ({\it dotted kernel}) or two kernels at
$x_i$ and $x_j$ ({\it left and right}) lead to the same summed estimate
at $x_s$. This shows a figure consisting of different types of
lines. Elements of the figure described in the caption should be set in
italics,
in parentheses, as shown in this sample caption. The last
sentence of a figure caption should generally end without a full stop}
\label{fig:example}
\end{figure}

If possible (e.g. if you use \LaTeX) please define figures as floating
objects. \LaTeX\ users, please avoid using the location
parameter ``h'' for ``here''. If you have to insert a pagebreak before a
figure, please ensure that the previous page is completely filled.

\subsection{Formulas}

Displayed equations or formulas are centered and set on a separate
line (with an extra line or halfline space above and below). Displayed
expressions should be numbered for reference. The numbers should be
consecutive within each section or within the contribution,
with numbers enclosed in parentheses and set on the right margin.
For example,
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt \;  .
\end{equation}

Please punctuate a displayed equation in the same way as ordinary
text but with a small space before the end punctuation.

\subsection{Program Code}

Program listings or program commands in the text are normally set in
typewriter font, e.g., CMTT10 or Courier.

\medskip

\noindent
{\it Example of a Computer Program}
\begin{verbatim}
program Inflation (Output)
  {Assuming annual inflation rates of 7%, 8%, and 10%,...
   years};
   const
     MaxYears = 10;
   var
     Year: 0..MaxYears;
     Factor1, Factor2, Factor3: Real;
   begin
     Year := 0;
     Factor1 := 1.0; Factor2 := 1.0; Factor3 := 1.0;
     WriteLn('Year  7% 8% 10%'); WriteLn;
     repeat
       Year := Year + 1;
       Factor1 := Factor1 * 1.07;
       Factor2 := Factor2 * 1.08;
       Factor3 := Factor3 * 1.10;
       WriteLn(Year:5,Factor1:7:3,Factor2:7:3,Factor3:7:3)
     until Year = MaxYears
end.
\end{verbatim}
%
\noindent
{\small (Example from Jensen K., Wirth N. (1991) Pascal user manual and
report. Springer, New York)}

\subsection{Footnotes}

The superscript numeral used to refer to a footnote appears in the text
either directly after the word to be discussed or -- in relation to a
phrase or a sentence -- following the punctuation sign (comma,
semicolon, or full stop). Footnotes should appear at the bottom of
the
normal text area, with a line of about 2~cm in \TeX\ and about 5~cm in
Word set
immediately above them.\footnote{The footnote numeral is set flush left
and the text follows with the usual word spacing. Second and subsequent
lines are indented. Footnotes should end with a full stop.}

\subsection{Citations}

The list of references is headed ``References" and is not assigned a
number
in the decimal system of headings. The list should be set in small print
and placed at the end of your contribution, in front of the appendix,
if one exists.
Please do not insert a pagebreak before the list of references if the
page is not completely filled.
An example is given at the
end of this information sheet. For citations in the text please use
square brackets and consecutive numbers: \cite{Alpher02},
\cite{Alpher03}, \cite{Alpher04} \dots

\bibliographystyle{splncs}
\bibliography{egbib}

\clearpage\mbox{}Page \thepage\ of the manuscript.
\clearpage\mbox{}Page \thepage\ of the manuscript.
\clearpage\mbox{}Page \thepage\ of the manuscript.
\clearpage\mbox{}Page \thepage\ of the manuscript.
\clearpage\mbox{}Page \thepage\ of the manuscript.
\clearpage\mbox{}Page \thepage\ of the manuscript.
This is the last page of the manuscript. 
\par\vfill\par
Now we have reached the maximum size of the ECCV 2010 submission.
\end{document}

### 2014-07-24 iccv2011/final.tex
move papers/* to top level

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{iccv}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{verbatim}
\usepackage{array}

\input{preamble}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pdftitle={Flint2011.pdf},pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

\iccvfinalcopy % *** Uncomment this line for the final submission

\def\iccvPaperID{1394} % *** Enter the ICCV Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ificcvfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Manhattan Scene Understanding Using Monocular, Stereo, and 3D Features}

\author{Alex Flint, David Murray, and Ian Reid\\
Active Vision Laboratory\\
Oxford University, UK\\
{\tt\small \{alexf,dwm,ian\}@robots.ox.ac.uk}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
}

\maketitle
\thispagestyle{empty}

%%%%%%%%% BODY TEXT
\input{main}

\end{document}

### 2014-07-24 iccv2011/final_addendum.tex
move papers/* to top level

--- 2014-07-24 iccv2011/final.tex
@@
-\documentclass[10pt,twocolumn,letterpaper]{article}
+\documentclass[10pt,letterpaper]{article}
 
@@
 \usepackage{array}
+\usepackage{xtab}
 
@@
 % run, let it finish, and you should be clear).
-\usepackage[pdftitle={Flint2011.pdf},pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
+\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
 
@@
 %%%%%%%%% TITLE
-\title{Manhattan Scene Understanding Using Monocular, Stereo, and 3D Features}
+\title{Addendum to Manhattan Scene Understanding Using Monocular, Stereo, and 3D Features}
 
@@
 %%%%%%%%% BODY TEXT
-\input{main}
+\input{main_addendum}
 

### 2014-07-24 iccv2011/main_addendum.tex
move papers/* to top level

\appendix

\section{Additional Results}

In this section we showcase further examples of models generated by
our system. In the table below, the left panel shows the base input
view $I_0$, the middle panel shows auxiliary views used for
photoconsistency calculations, and the right panel shows the MAP model
$\Model$ inferred by our system. In all our experiments we used two
image auxiliary images per base image, which were sampled one second
before and one second after the base image in the video sequence. For
further details of parameter settings see the main paper.

\newcommand{\ResultIm}[7]{\includegraphics[width=#1\textwidth]{figures_addendum/#2/#3_#4_frame#5_#6.#7}}
\newcommand{\ShowcaseIm}[4]{\ResultIm{0.23}{resized_full_results}{#1}{#2}{#3}{#4}{jpg}}
\newcommand{\AuxIm}[4]{\ResultIm{0.1}{aux_links/resized}{#1}{#2}{#3}{#4}{jpg}}
\newcommand{\ShowcaseRow}[3]{
        \ShowcaseIm{#1}{#2}{#3}{orig} &
        \AuxIm{#1}{#2}{#3}{aux0}
        \AuxIm{#1}{#2}{#3}{aux1} &
        \ShowcaseIm{#1}{#2}{#3}{dp} &
}
\newcommand{\ShowcaseRowLast}[3]{  % contains no final auxiliary image
        \ShowcaseIm{#1}{#2}{#3}{orig} &
        \AuxIm{#1}{#2}{#3}{aux0}
        \AuxIm{#1}{#2}{#3}{aux0} &
        \ShowcaseIm{#1}{#2}{#3}{dp} &
}
\newcommand{\ShowcaseRowFirst}[3]{  % contains no previous auxiliary image
        \ShowcaseIm{#1}{#2}{#3}{orig} &
        \AuxIm{#1}{#2}{#3}{aux1}
        \AuxIm{#1}{#2}{#3}{aux1} &
        \ShowcaseIm{#1}{#2}{#3}{dp} &
}

\newcommand\ColHeadings{\textbf{Input (base view)} & \textbf{(auxiliary views)} & \textbf{Output of our system}}

\begin{center}
\begin{tabular}{cccc}
  \ColHeadings \\
  \ShowcaseRow{exeter}{bursary}{006} \\
  \ShowcaseRow{exeter}{bursary}{026} \\
  \ShowcaseRow{exeter}{bursary}{028} \\
  \ShowcaseRow{exeter}{mcr1}{012} \\
\end{tabular}

%\clearpage
\begin{tabular}{cccc}
  \ColHeadings \\
  \ShowcaseRow{exeter}{mcr1}{014} \\
  \ShowcaseRow{exeter}{mcr1}{015} \\
  \ShowcaseRow{exeter}{mcr1}{019} \\
  \ShowcaseRow{exeter}{mcr1}{025} \\
  \ShowcaseRow{exeter}{mcr1}{044} \\
  \ShowcaseRow{exeter}{mcr1}{051} \\
  \ShowcaseRowLast{exeter}{mcr1}{054} \\
\end{tabular}

\clearpage
\begin{tabular}{cccc}
  \ColHeadings \\
  \ShowcaseRow{lab}{atrium2}{009} \\
  \ShowcaseRow{lab}{foyer1}{008} \\
  \ShowcaseRow{lab}{foyer1}{015} \\
  \ShowcaseRow{lab}{foyer1}{015} \\
  \ShowcaseRow{lab}{foyer1}{021} \\
  \ShowcaseRow{lab}{foyer1}{040} \\
  \ShowcaseRow{lab}{foyer2}{002} \\
\end{tabular}

\clearpage
\begin{tabular}{cccc}
  \ColHeadings \\
  \ShowcaseRow{lab}{foyer2}{006} \\
  \ShowcaseRow{lab}{foyer2}{014} \\
  \ShowcaseRow{lab}{foyer2}{017} \\
  \ShowcaseRow{lab}{foyer2}{036} \\
  \ShowcaseRow{lab}{foyer2}{041} \\
  \ShowcaseRow{lab}{foyer2}{042} \\
  \ShowcaseRow{lab}{ground1}{008} \\
\end{tabular}

\clearpage
\begin{tabular}{cccc}
  \ColHeadings \\
  \ShowcaseRow{lab}{ground1}{011} \\
  \ShowcaseRow{lab}{ground1}{020} \\
  \ShowcaseRow{lab}{ground1}{025} \\
  \ShowcaseRow{lab}{ground1}{032} \\
  \ShowcaseRow{lab}{ground1}{039} \\
  \ShowcaseRow{lab}{kitchen1}{004} \\
  \ShowcaseRow{lab}{kitchen1}{030} \\
\end{tabular}

\clearpage
\begin{tabular}{cccc}
  \ColHeadings \\
  \ShowcaseRow{lab}{kitchen1}{044} \\
  \ShowcaseRow{lab}{kitchen1}{078} \\
  \ShowcaseRowFirst{som}{corr1}{001} \\
  \ShowcaseRow{som}{corr1}{012} \\
  \ShowcaseRow{som}{corr1}{015} \\
  \ShowcaseRow{som}{corr1}{018} \\
  \ShowcaseRow{som}{corr1}{020} \\
\end{tabular}

\end{center}

\section{Payoffs}

In this section we provide some visualizations of the payoff matrices
discussed in the main paper. We hope these visualizations give extra
insights into the strenghts and weaknesses of each sensor modality.

\newcommand\FooPayoffImg[2]{
        \parbox[c]{1em}{
                \includegraphics[width=0.3\textwidth]{figures_addendum/lab_foyer2_frame#1_#2.png}}}
\newcommand\PayoffImg[1]{\FooPayoffImg{010}{#1}}

\newcolumntype{I}{>{\arraybackslash} m{.3\linewidth} }
\newcolumntype{N}{>{\arraybackslash} m{.3\linewidth} }
\newcolumntype{T}{>{\arraybackslash} m{.6\linewidth} }

%\begin{tabular}{lp{0.6\textwidth}}
\begin{tabular}{IT}
  \PayoffImg{orig} &
  The raw image provided as input to our system. \\
\end{tabular}

\begin{tabular}{IT}
  \PayoffImg{gt} &
  The ground truth segmentation for this image. Horizontal surfaces
  are shaded blue. Vertical surfaces are shaded red and green. We will
  refer to these as the ``red'' and ``green'' orientations.\\
\end{tabular}

\begin{tabular}{IT}
  \PayoffImg{dp} &
  The MAP indoor manhattan model $\Model$ output by our system for
  this input. \\
\end{tabular}

\begin{tabular}{IT}
  \PayoffImg{monopayoffs0} &
  Payoffs $\MonoPayoff$ derived from monocular image features, for the
  ``green'' orientation. Pixels of higher intensity correspond to
  larger values in the payoff matrix. The MAP model is shown in
  wireframe using red lines. Intuitively, the optimization over models
  can be thought of as finding the minimal cost path through the
  payoff matrix, where higher intensity pixels correspond to lower
  costs. This is only a rough picture, however; the real optimization
  situation is more complex since models are penalized for each
  additional corner. \\
\end{tabular}

\begin{tabular}{IT}
  \PayoffImg{monopayoffs1} &
  As above, for the ``red'' surface orientation. \\
\end{tabular}

\begin{tabular}{IIN}
  \FooPayoffImg{009}{orig} &
  \FooPayoffImg{011}{orig} &
  Auxiliary images used for stereo photoconsistency. In our
  experiments we used two image auxiliary images for each base image,
  which were sampled one second before and one second after the base
  image in the video sequence.
\end{tabular}

\begin{tabular}{IIN}
  \PayoffImg{stereopayoffs_aux0} &
  \PayoffImg{stereopayoffs_aux1} &
  Payoffs $\StereoPayoff$ corresponding to the auxiliary images
  above. Each pixel represents the photoconsistency score for a wall
  segment with floor/wall (or ceiling/wall) intersection $\ModelAtX$
  at that pixel. Notice the repeated ``pizza slice'' patterns in which
  one tip of the triangle is located at the floor/wall intersection. \\
\end{tabular}

\begin{tabular}{IT}
  \PayoffImg{points} &
  Here we show the structure--from--motion point cloud. The points are
  shown projected into the image, but the system has access to their
  3D locations. Notice how the points are not uniformly distributed in
  the image.
\end{tabular}

\begin{tabular}{IT}
  \PayoffImg{projs} &
  Here we show the structure--from--motion point cloud projected onto
  the floor and ceiling planes, which were recovered as a seperate
  step as described in the main paper. The red dots show the original
  3D point cloud and the blue dots show the projections onto the floor
  and ceiling.\\
\end{tabular}

\begin{tabular}{IT}
  \PayoffImg{3dpayoffs_agree} &
  This shows the component of the payoffs $\DepthPayoff$ intended to
  provide a bias towards models that explain the observed 3D
  points. This is the component corresponding to $t=\ON$. Each bright
  spot corresponds to the projection of a 3D point onto the floor or
  ceiling plane.
\end{tabular}

\begin{tabular}{IT}
  \PayoffImg{3dpayoffs_occl} &
  This shows the component of the payoffs $\DepthPayoff$ intended to
  penalize walls that occlude observed 3D points. This corresponds to
  the case that $t\in\{\IN,\OUT\}$. Notice that for each 3D point,
  payoffs are assigned to walls that pass between the floor and
  ceiling projection of that point. Such walls are precisely those
  which do \textit{not} occlude the point.\\
\end{tabular}

\begin{tabular}{IT}
  \PayoffImg{payoffs0} &
  Joint payoff matrix $\JointPayoff$. \\
\end{tabular}

### 2014-07-24 techreport_oct2010/main.tex
move papers/* to top level

\documentclass{article}

\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{titlesec}
\usepackage{algorithm,algorithmic}
\usepackage{subfig}
\usepackage{latexsym}

\input{../common/AlexDefs}

\newcommand\Hcf{H_{c \rightarrow f}}
\newcommand\Hcfhat{\hat{H}_{c \rightarrow f}}

\newcommand\vvpt{\vect{v_v}}
\newcommand\lvpt{\vect{v_l}}
\newcommand\rvpt{\vect{v_r}}

\newcommand{\abs}[1]{\lvert#1\rvert}
\newcommand{\norm}[1]{\lVert#1\rVert}

\newcommand\Payoff{\Pi}
\newcommand\ColPayoff{\pi}
\newcommand\Model{M}
\newcommand\ManhattanClass{\mathcal{M}}
\newcommand\Feature{\vect{\psi}}
\newcommand\Features{\Psi}
\newcommand\Pixel{\vect{p}}
\newcommand\Pixels{\vect{P}}
\newcommand\Orient{a}
\newcommand\Depth{d}
\newcommand\Depths{D}
\newcommand\Ind{t}

\newcommand\Penalties{\vect{\lambda}}
\newcommand\PenaltyOccl{\lambda_\textsf{occl}}
\newcommand\PenaltyConc{\lambda_\textsf{conc}}
\newcommand\PenaltyConv{\lambda_\textsf{conv}}

\newcommand\tIN{\tau_{\textsf{IN}}}
\newcommand\tOUT{\tau_{\textsf{OUT}}}
\newcommand\tON{\tau_{\textsf{ON}}}

\newcommand\fIN{f_{\textsf{in}}}
\newcommand\fOUT{f_{\textsf{out}}}
\newcommand\fUP{f_{\textsf{up}}}
\newcommand\fDOWN{f_{\textsf{down}}}

\input GotIn.fd
\newcommand*\initfamily{\usefont{U}{GotIn}{xl}{n}}

\begin{document}

\section{Abstract}

\section{Introduction}

\initfamily
\fontsize{50}{60}
\selectfont
O\fontsize{10}{12}\normalfont ver the past decade, computer vision researchers working with
monocular images have pursued substantially different research
directions to those working with multiple views. The focus for
monocular images has increasingly been to infer high--level facts
about the world, such as the locations of and interactions between
objects, to semantic scene categories and the spatial layout of the
environment. In constrast, much of the work concerning multiple views
has focussed on reconstructing camera poses and scene structure using
techniques such as structure--from--motion, stereo, and multiple--view
stereo, with performance metrics typically concerned with geometric
accuracy of the reconstructions.

In this paper we focus on utilizing multiple view geometry for image
understanding purposes. We assume a moving camera with a
structure--from--motion system estimating its trajectory through
space, and show how to infer semantically meaningful models of the
environment. We focus on the \textit{indoor Manhattan
  representation}\cite{Lee09}, in which the world is modelled
in terms of floor, wall, and ceiling surfaces. This representation
captures many semantically meaningful aspects of the environment,
including
\begin{enumerate}
  \item{\textbf{Scale}: the distance from floor to ceiling suggests
    a scale for distances in the environment}
  \item{\textbf{Boundaries}: walls constrain movement in the
    environment and suggest locations for doors, windows, and other
    objects.}
  \item{\textbf{Gravity}: the orientation of the ground plane implies
    the direction in which gravity operates, which constrains the
    arrangement of objects resting upon one another.}
  \item{\textbf{Shape}: the organisation of walls in an environment
    suggests its functional category (kitchen/office/etc).}
\end{enumerate}

Indoor manhattan models are useful because they capture these
properties explicitly as primitives of the model, whereas to extract
such properties from a dense polygonal mesh reconstruction would
require additional non--trivial inference following the reconstruction
stage. Our approach can be seen as inferring semantic properties of
the scene directly from multiple--view data, without an intermediate
dense reconstruction step. This makes sense if, as is the case in our
context, these proprties constitute the ultimate goal of the
system. Of course, if a photo--realistic reconstruction is itself the
end goal then our approach is not suitable.

We build on the dynamic programming approach to inferring indoor
manhattan models of Flint \etal \cite{FlintECCV10}. Our work extends
their approach to allow joint inference across multiple views as well
as to incorporate range data alongside photometric cues. This involves
a significant generalization of their approach since neither of these
new modalities are representable using their pixel--wise cost
function. We therefore generalize their approach to a payoff--based
formulation permitting more general costs. We also show that in this
new formulation one of the state variables involved in their
optimisation is unecessary; its removal corresponds to eliminating an
entire dimension from the search and results in an order of magnitude
efficiency improvement.

XXX Probabilistic cost function

XXX Combine monocular, multi-view, and range features

When comparing our approach to previous work we propose that in
addition to measuring geometric accuracy, the \textit{size} of
alternative representations should also be compared, since models that
capture key properties of the environment using fewer representational
bits must have done a better job of filtering salient information
from sensor data. This is in line with the relationship between the
ability of an algorithm to compress data and its ability to make
predictions (cite Hutter).

The remainder of this paper is organised as follows. XXX

\section{Background}

The Manhattan world assumption has seen increasing attention in the
computer vision literature over past years
\cite{Coughlan99,Zhang02,Lee09,Furukawa09,FlintECCV10}. It was
introduced by Coughlan and Yuille\cite{Coughlan99} for estimating
camera orientations. Furukawa \etal \cite{Furukawa09} proposed a
``Manhattan--world stereo'' algorithm that incorporates Manhattan
orientations into an energy function, which they solve using graph
cuts.

Brostow et al

Talk more about multi view reconstructions

BUILDING RECONSTRUCTION USING MANHATTAN WORLD GRAMMARS

The ECCV(?) paper on using manhattan + non-manhattan surfaces for
reconstruction.

While their approach is concerned with dense photo--realistic
reconstructions, ours is intended to capture semantic properties of
the scene using a representation that is as concise as possible.

Lee \etal \cite{Lee09} first proposed indoor Manhattan models (a
sub--class of general Manhattan models) for monocular
reconstructions. They used a branch--and--bound algorithm together
with a line--sweep heuristic for approximate inference. Flint \etal
\cite{Flint10ECCV} inferred similar models using a dynamic programming
algorithm. In earlier work \cite{Flint10CVPR} they demonstrated
Manhattan reconstructions integrated with a SLAM system, but this work
inferred models from single frames and then extrapolated these forward
in time. In contrast, the present work allows both photoconsistency
and depth information to be included directly in the cost function,
allowing far more robust inference.

XXX We also use features and learning rather than a line sweep heuristic...

Felzenszwalb and Veksler \cite{Felzenszwalb10} posed the
reconstruction problem in terms of an energy minimisation, which they
showed could be solved using dynamic programming. Barinova \etal
\cite{Barinova08} proposed to model outdoor scenes using a conditional
random field over vertical plane boundaries. However, neither of these
approaches capture the geometry of indoor Manhattan scenes in the
strong sense, and neither of them optimize over cost functions that
can incorporate multiple views or depth data.

\section{Model}

\begin{figure}[tb]%
  \centering
  \label{fig:ideal-models}
  \begin{tabular}{ccc}
    \includegraphics[width=0.2\textwidth]{figures/true_models/lab_ground1_010_rect.png} &
    \includegraphics[width=0.2\textwidth]{figures/true_models/lab_kitchen_030_rect.png} &
    \includegraphics[width=0.2\textwidth]{figures/true_models/lab_ground1_030_rect.png} \\
    \includegraphics[width=0.2\textwidth]{figures/true_models/lab_ground1_010} &
    \includegraphics[width=0.2\textwidth]{figures/true_models/lab_kitchen_030} &
    \includegraphics[width=0.2\textwidth]{figures/true_models/lab_ground1_030}
  \end{tabular}
  \hspace{0.2cm}
  \caption{Three input images and the indoor Manhattan models we
    seek. Notice how each image column intersects exactly one
    wall.}
\end{figure}

In this section we describe the class of indoor Manhattan models and
pose inference within this class as maximization of a payoff function
that decomposes over image columns. In subsequent sections we will
show that reasonable generative models for photometric,
multiple--view, and depth features give rise to MAP inference formulae
that decompose in precisely this way, so that, given appropriate
mapping of sensor inputs to payoffs, maximization of the payoff
function is equivalent to MAP inference in a graphical model. This
will allow us in \sectref{inference} to provide a concise dynamic
programming solution in the setting of payoff matrices.

An indoor Manhattan model consists of a floor plane, a parallel
ceiling plane, and a set of vertical walls extending between
them. Each wall extends all the way from the floor to ceiling, and
pairs of walls meet at vertical edges. Walls are restricted to two
orthogonal orientations, and the camera is assumed to be located
between the floor and ceiling planes.

We always consider indoor Manhattan models as seen from a particular
camera. In can be shown that, due to the assumption of walls extending
all the way from floor to ceiling, indoor Manhattan models always
project as a sequence of walls that only ever meet (in the image) their
left and right neighbours (in the sequence).

We assume that vanishing points for each of the three Manhattan
directions are given. There are several well--known techniques for
extracting vanishing points from images\cite{Shufelt99}; for the
monocular case we use the EM approach of \cite{Zhang02}, and in the
multiple view setting we use the joint estimation approach of
\cite{FlintCVPR10}.

It will simplify the remainder of this paper if we assume that
vertical lines in the world appear vertical in images. Given the
location of the vertical vanishing point $\vvpt$ we rectify images
according to the homography
\begin{equation}
  H = 
  \begin{pmatrix}
    \vvpt \times \vect{e_3} \\
    \vvpt \\
    \vvpt \times \vect{e_3} \times \vvpt
  \end{pmatrix}
  ~,~~~~~~~~ \vect{e_3}=[0,0,1]^T ~.
\end{equation}
This will introduce warping artifacts for cameras with a
near--vertical principal ray; however, this rectification step is not
strictly necessary and we could instead perform geometric calculations
in the original image coordinates, though doing so would signficantly
complicate the exposition of this paper.

Following rectification in can be shown that each image column
contains exactly one wall segment; for example, see
\figref{ideal-models}. Let the position of the floor--wall boundary in
image column $x$ be $y_f(x)$, and the position of the corresponding
ceiling--wall boundary be $y_c(x)$. We define the payoff
$\Payoff(\Model)$ for the model $\Model$ as
\begin{equation}
  \label{eq:abstract-payoffs}
  \Payoff(\Model) = \sum_x \Bigl(\ColPayoff(x,y_f) +
  \ColPayoff(x,y_c)\Bigr) ~.
\end{equation}
where $a$ is a payoff matrix assigning to each pixel a payoff for
models that label that pixel as a floor-wall or ceiling--wall
intersection. Note that $a(x,y)$ is \textit{not} restricted to
dependence on pixel $(x,y)$, or even to a local region about $(x,y)$,
and indeed the payoff functions described in the next section
incorporate image evidence from image regions far from $(x,y)$.

Where to put this?? Since the 3D points corresponding to $y_c$ and
$y_f$ are vertically above/beneath one another, and because the floor
and ceiling planes are parallel to one another, $y_c$ and $y_f$ are
related to one another via a planar homology\cite{Criminisi01} $G$,
\begin{equation}
  \begin{pmatrix} x \\ y_c \\ 1 \end{pmatrix}
  \hspace{0.3cm} \sim \hspace{0.3cm}
  G \begin{pmatrix} x \\ y_f \\ 1 \end{pmatrix}
  ~.
\end{equation}

TODO: how to incorporate penalties for additional walls etc.

\subsection{Monocular features}

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/monocular-gm}
  \caption{The graphical model relating building structures $M$ to
    monocular image features $\Feature$. $\Pixel=(x,y)$ is an observed
    pixel location and $a$ is indicates the orientation of the model
    $M$ at that pixel.}
  \label{fig:monocular-gm}
\end{figure}

To infer indoor Manhattan models from monocular images we assume the
graphical model shown in \figref{monocular-gm}. Pixel features
$\Feature$ are dependent on the orientation $\Orient\in\{1,2,3\}$ at
that pixel, where the possible values for $\Orient$ correspond to the
three manhattan orientations. We assume a linear model for pixel
features,
\begin{equation}
\label{eq:feature-likelihood}
  P(\Feature ~|~ \Orient) =
  \frac{{\vect{\theta}_\Orient}^T\Feature}
       {\sum_j{\vect{\theta}_\Orient}^T\Feature_j} ~.
\end{equation}

A model $\Model$ generates pixel orientations $\Orient$
deterministically. We include this variable for notational
convenience only, thus
\begin{equation}
  \label{eq:orient-model}
  P(\Orient ~|~ \Model,\Pixel) = 
    \begin{cases}
      1, & \mbox{if } \Orient = a^*(\Model,\Pixel) \\
      0, & \mbox{otherwise}\\
    \end{cases}
    ~,
\end{equation}
where $a^*(\Model,\Pixel)$ denotes the orientation predicted by model
$\Model$ at pixel $\Pixel$, which we compute by filling polygons in
the image. Finally, for a model with $n_1$ concave corners, $n_2$
convex corners, and $n_3$ occluding corners, we assume the complexity
prior
\begin{eqnarray}
  \label{eq:model-prior}
  P(\Model ~|~ \Penalties) &=& \frac{1}{Z} 
    \PenaltyConc^{n_1} \PenaltyConv^{n_2} \PenaltyOccl^{n_3} \\
  Z &=& \sum_{n_c,n_n,n_o} P(\Model ~|~ \Penalties) \\
    &=& \frac{1}{(1-\PenaltyConc)(1-\PenaltyConv)(1-\PenaltyOccl)}
\end{eqnarray}
This corresponds to a fixed probability for ``events'' corresponding
to each type of corner. XXX need to introduce the three types of
corners earlier.

We now derive MAP inference for indoor manhattan models. The posterior
on building structures given observed image features is
\begin{eqnarray}
  P(\Model ~|~ \Features,\Pixels) &=&
  \eta P(\Model) P(\Features,\Pixels ~|~ \Model)
  \\
  &=& \eta P(\Model) \prod_i P(\Feature_i,\Pixel_i ~|~ \Model)
  \\
  &=& \eta \sum_A P(\Model) \prod_i
    P(\Feature_i,\Pixel_i, \Orient_i  ~|~ \Model)
  \\
  &=& \eta \sum_A P(\Model) \prod_i P(\Feature_i ~|~ \Orient_i)
    P(\Orient_i ~|~ \Pixel_i,\Model) ~.
\end{eqnarray}
But according to \eqnref{orient-model}, the only non--zero term in the
summation over $A=\{a_i\}_{i=0}^N$ is the one for which $a_i =
  a_i^*(\Model,\Pixel_i)$. Hence,
\begin{eqnarray}
  \label{eq:mono-post}
  P(\Model ~|~ \Features,\Pixels) &=&
  \eta P(\Model) \prod_i P(\Feature_i ~|~ \Orient_i^*) \\
  \label{eq:mono-logpost}
  \log P(\Model ~|~ \Features,\Pixels) &=&
  c + n_1 \log \PenaltyConc + n_2 \log \PenaltyConv + n_o \log \PenaltyOccl \\
  && + \sum_i \log P(\Feature_i ~|~ \Orient_i^*)
\end{eqnarray}
where $c$ corresponds to the normalizing denominators in
\eqnref{mono-post} and \eqnref{model-prior}.

The sumation in \eqnref{mono-logpost} is over pixel coordinates, but
form required by \eqnref{abstract-payoffs} is over image columns. In
order that the maximum--payoff model coincide with the MAP solution
to \label{eq:mono-logpost} we set
\begin{equation}
  \ColPayoff(x,y) = \sum_{y'} \log P(\Feature_i ~|~ \Orient_{y'}^*(y))
\end{equation}
where $\Orient_{y'}^*(y)$ is the orientation for pixel $(x,y')$
generated by a model with a floor/wall or ceiling/wall boundary at
$(x,y)$. Since we assume that vertical lines in the world project to
vertical image lines, this mapping is independent of $x$.

XXX expand on this, give some visualizations.

\subsection{Multiple view features}

We now formulate the payoff matrix $\ColPayoff(x,y)$ for the case
that multiple views of the scene are available. We assume that one
frame is selected as the base frame $I_0$; the remaining $N\geq1$
auxiliary frames are denoted $I_1,\ldots,I_N$. We assume poses
$P_i=\{R_i,\vect{t_i}\}$ are given, for example as output by a
structure--from--motion system. We further assume that camera
calibrations $C_i$ are known, though different cameras may have
different calibrations.

Intuitively, we treat inference in this settings as follows: we
consider models $M$ in terms of their projection into the base frame
$I_0$, exactly as in the previous section. Since edges are explicitly
associated with vanishing points, a hypothesized model $M$ in image
coordinates uniquely specifies a full 3D model up to scale (which we
show how to resolve later). Because camera poses are known we can
back--project any pixel from the base frame onto the 3D model then
re--project it into the auxiliary frames, giving a pixel--wise
correspondence between the base and auxiliary frames as shown in
\figref{backproject}. We then optimise models with respect to the
photoconsistency of the pixel correspondences they
introduce. Cruicially, the photoconsistency function decomposes
linearly over image columns, and the contribution of any image column
$x$ is dependent on the model $M$ only through the location of the
floor/wall intersection $y_f(x)$ at that column. XXX modulo
occlusions! This is a result of the specific structure of indoor
Manhattan models and allows fast, exact, and deterministic MAP
inference.

\begin{figure}[tb]
  \centering \includegraphics[width=0.4\textwidth]{figures/backproject}
  \caption{Pixel correspondences across multiple views are computed by
    back--projection into the world followed by re--projection in
    auxiliary views.}
  \label{fig:backproject}
\end{figure}

Need to show how to get $z$ coords of the planes.

\subsubsection{Extension: using orientation information}
Find floor and ceiling by maximizing NCC (weighted at bottom of
image?) w.r.t. floor and ceiling positions?

\subsubsection{Extension: using orientation information}

\subsubsection{EM for Occlusion Resolution}

\subsection{Depth features}

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/3d-gm}
  \caption{The graphical model relating indoor Manhattan models to 3D
    points. The hidden variable $t$ indicates whether the point is
    inside, outside, or coincident with the model.}
  \label{fig:monocular-gm}
\end{figure}

In this section we explore the context in which a 3D point cloud, as
generated perhaps by structure--from--motion or depth--sensing
hardware, is available to the system. For notational convenience we
assume that points are represented as pixel/depth pairs
$X=\{\Pixel_i,\Depth_i\}_{i=0}^N$. This representation is trivially
computed from a point cloud. We assume the following generative
model. First, an image point $\Pixel_i$ is sampled at random, then a
depth measurement $d_i$ is performed along the corresponding ray. The
relationship between $d_i$ and the model $M$ is determined by a hidden
variable $t_i\in\{\textsf{IN},\textsf{OUT},\textsf{ON}\}$ as follows:
\begin{enumerate}
  \item{If $t_i=\textsf{IN}$ then $\Depth_i$ corresponds to an object
    or surface within the room.}
  \item{If $t_i=\textsf{OUT}$ then $\Depth_i$ corresponds to an object
    or surface outside the boundary of the room, perhaps observed
    through a window or doorway. This case can also account for
    outliers generated at an earlier stage of processing.}
  \item{If $t_i=\textsf{ON}$ then $\Depth_i$ corresponds to the
    closest surface in $M$ in the direction of $\Pixel_i$.}
\end{enumerate}
We assign a fixed prior to $t$,
\begin{eqnarray}
  P(t=\textsf{IN}) = \tIN \hspace{0.5cm} 
  P(t=\textsf{OUT}) = \tOUT \hspace{0.5cm} 
  P(t=\textsf{ON}) = \tON ~.
\end{eqnarray}
We assume that each measurement is constrained to a known interval
$\left[0,d_{\textsf{max}}\right]$. Let $d_\Model(\Pixel)$ be the
depth to the nearest surface of model $M$ in the direction of
$\Pixel$ (see fig XXX). We model depth measurements accordinging to
\begin{eqnarray}
  \label{eq:x-inside}
  P(d~|~\Model,\Pixel,t=\textsf{IN}) &=&
  \begin{cases}
    \frac{1}{d_\Model(\Pixel)}, & \mbox{if } d_i < d_\Model(\Pixel) \\
    0, & \mbox{otherwise} \\
  \end{cases}\\
  \label{eq:x-outside}
  P(d~|~\Model,\Pixel,t=\textsf{OUT}) &=&
  \begin{cases}
    \frac{1}{d_{max}-d_\Model(\Pixel)}, & \mbox{if } d_i > d_\Model(\Pixel) \\
    0, & \mbox{otherwise} \\
  \end{cases}\\
  P(d ~|~ \Model,\Pixel,t=\textsf{ON}) &=&
  \mathcal{N}(d~;~d_\Model(\Pixel),\sigma)
\end{eqnarray}

\begin{figure}[tb]
  \centering \includegraphics[width=0.4\textwidth]{figures/point-projs}
  \caption{Three possibilities for depth measurements: inside, outside, and
    coincident.}
  \label{fig:backproject}
\end{figure}

The posterior on $M$ is
\begin{eqnarray}
  \label{eq:3d-posterior}
  P(\Model~|~\Depths,\Pixels) &=& P(\Model) \prod_i P(\Depth_i,\Pixel_i~|~\Model) \\
  P(\Depth_i,\Pixel_i~|~\Model) &=& \sum_{t_i} P(\Depth_i,\Pixel_i,t_i~|~\Model) \\
  &=& \sum_{t_i} P(\Depth_i,~|~\Pixel_i,t_i,\Model) P(t_i~|~\Model) ~,
\end{eqnarray}
where in the second line we have omitted the prior on $\Pixel_i$ as it
is uniform over the image and will play no part in the maximisation to
come. Now \eqnref{3d-full-post} is a sum over three possible values
for $t_i$ and we can simplify this expression by distinguishing the
case where $d_i>d_\Model(\Pixel)$, in which case \eqnref{x-inside}
is $0$, from the case where the opposite is true, in which case
\eqnref{x-outside} is $0$. Hence,
\begin{eqnarray}
  \label{eq:3d-likelihood}
  P(\Depth_i~|~\Pixel_i,\Model) &=& \alpha +
  \tON \mathcal{N}(d_o~;~d_\Model(\Pixel),\sigma) \\
  \alpha &=&
  \begin{cases}
    \frac{\tIN}{d_\Model(\Pixel)}, & \mbox{if } d_i < d_\Model(\Pixel_i) \\
    \frac{\tOUT}{d_{max}-d_\Model(\Pixel)}, & \mbox{otherwise} \\
  \end{cases}~.
\end{eqnarray}

Notice that if the floor/wall intersection of a model $M$ meets column
$x$ at $y=y_f$, then this alone is sufficient to compute the depth
$d_(y';x,y_f)$ of the model at each pixel $(x,y')$ in that column. We
use this to express the posterior \eqnref{3d-posterior} in the payoff
form required by \eqnref{abstract-payoff}. Let $\mathcal{C}_x$ contain
indices for all depth measurements in column $x$, then
\begin{eqnarray}
  P(\Model ~|~ \Depths,\Pixels) &=& 
  \prod_{x} \prod_{i\in\mathcal{C}_x} P(\Depth_i~|~\Pixel_i,\Model) \\
  \log P(\Model ~|~ \Depths,\Pixels) &=& 
  P(M) + \sum_{x} \Bigl( \sum_{i\in\mathcal{C}_x}
  \log P(\Depth_i~|~\Pixel_i,y_f(x)) \Bigr) ~,
\end{eqnarray}
which we write in payoff form as
\begin{equation}
  \ColPayoff(x,y) = 
  \sum_{i\in\mathcal{C}_x} \log P(\Depth_i~|~\Pixel_i,y)
\end{equation}

TODO: explain where the prior on models went.

\subsection{Joint distribution}

We can trivially combine photometric, stereo, and depth data into a
joint model by assuming conditional independence between the sensor
modalities given $M$,
\begin{equation}
  P(M ~|~ X_{\textsf{mono}}, X_{\textsf{stereo}}, X_{\textsf{3D}})
  =  P(M) P(X_{\textsf{mono}} ~|~ M)
          P(X_{\textsf{stereo}} ~|~ M)
          P(X_{\textsf{3D}} ~|~ M) ~.
\end{equation}
Taking logarithms leads to simply summing over payoffs,
\begin{equation}
  \ColPayoff_{\textsf{joint}}(\vect{x}) =
  \ColPayoff_{\textsf{mono}}(\vect{x}) + 
  \ColPayoff_{\textsf{stereo}}(\vect{x}) +
  \ColPayoff_{\textsf{3D}}(\vect{x}) ~.
\end{equation}

\section{Inference}

\newcommand\OptModel{\Model^*}
\newcommand\PartialModel{\Model}
\newcommand\CroppedModel{\Model'}
\newcommand\PartialCol{x}
\newcommand\CroppedCol{x'}

We have reduced MAP inference to optimization over a payoff matrix:
\begin{equation}
  \OptModel = \argmax_\Model\limits \sum_x \ColPayoff\Bigl(x,y_f(\Model)\Bigr)
\end{equation}
Flint \etal \cite{FlintECCV10} showed that if an indoor Manhattan
model $\PartialModel$ is optimal over image columns
$\left[0,\PartialCol\right]$, then the ``cropped'' model
$\CroppedModel$, obtained by restricting $\PartialModel$ to the
sub--interval $\left[1,\CroppedCol\right]$, $\CroppedCol<\PartialCol$,
must be optimal over $\left[1,\CroppedCol\right]$. This permits a
dynamic programming solution in which the MAP model $\OptModel$ is
obtained by building up models from left to right.

Our algorithm differs from that of \cite{FlintECCV10} in that we do
not include the number of corners in the model as a state variable,
but instead accumulate the penalties $\Penalties$ each time we
consider adding a corner. This reduces complexity by $O(K)$, where $K$
is the number of walls in the model. We also operate directly on the
payoff matrix $\ColPayoff$, rather than by summing over pixel
agreements as in previous work. This is the key insight that allows 3D
and stereo data to be incorporated into the optimisation. For
completeness, we give the new recurrence relations below, though for
the full treatment the reader is referred to \cite{FlintECCV10}.
\begin{eqnarray}
  \label{eq:fout-recurrence}
    \fOUT(x,y,a) &=& \min_{a'\in\{l,r\}} \min
      \begin{cases}
        \fUP(x,y-1,a') + \PenaltyConc & \\
        \fDOWN(x,y+1,a') + \PenaltyOccl & \\
        \fIN(x,y,a') + \PenaltyOccl &
      \end{cases}\\
  \label{eq:fup-recurrence}
  \fUP(x,y,a) &=& \min \Bigl(\fIN(\cdot), \fUP(x,y-1,a)\Bigr)~,\\
  \label{eq:fdown-recurrence}
  \fDOWN(x,y,a) &=& \min \Bigl(\fIN(\cdot), \fDOWN(x,y+1,a)\Bigr)~,\\
  \label{eq:fin-recurrence-basic}
  \fIN(x,y,a) &=&
    \min_{x'<x}\Bigl(\fOUT(x',y',a)+ \Delta\Bigr) ~,\\
  \Delta &=& - \sum_{i=x'}^{x} \ColPayoff(i,y')\\
  \fOUT(0,y,a) &=& 0 \hspace{1cm} \forall y,a \\
  \fUP(x,0,a) &=& \infty \hspace{1cm} \forall x,a \\
  \fDOWN(x,N_x,a) &=& \infty \hspace{1cm} \forall x,a
\end{eqnarray}
XXX: These are useless, maybe just omit the recurrence relations? Or
move to appendix?

\section{Training}

Parameters to train from data
\begin{enumerate}
  \item{the penalties $\vect{\lambda}$ for the prior $P(M)$}
  \item{for monocular: the linear classifiers $\theta$}
  \item{for stereo: the photoconsistency metric $\Sigma$}
  \item{for 3D: the parameters $\vect{\tau}$ and the Gaussian
    variance $\sigma$.}
\end{enumerate}

\section{Results}

\bibliographystyle{splncs}
\bibliography{AVLstrings,VisionRefs}

\end{document}
